---
title: "Cleaning_Production-data_MMO"
author: "AL"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
# sets options for entire document in the first chunk
knitr::opts_chunk$set(echo = TRUE)
```

## Preparation

```{r Load libraries, include=FALSE}
# Load packages
library(tidyr) # for tidying data
library(dplyr) # base R alternative but neater
library(vroom) # for loading and transforming data
library(tidyverse) # data exploration 
library(reshape2) # for melting data frames i.e. short and wide to long and thin
library(data.table) # for fread()-function
library(mice) # md.pattern to show missing data
library(stringr) # used to replace matched patterns in a string
library(haven) #Read SAV files
library(readxl)
library(readr) # CSV file I/O, e.g. the read_csv function
library(sjlabelled)
library(compareDF)#Check missing values across datasets
library(chron)

```


```{r Load data, echo=TRUE}
# Loading file path to the data into a Volume 
source("Data_filepath.R")# Data_filepath.R is listed in .gitignore-file. So, you will need to create that file yourself and provide your respected filepath using "data_dir <- [enter your here]"
```


```{r Specify the desired file path}
filepath <- paste(data_dir,"Methods/SpeciesTypeClassification", sep = "") 
# note we are able to change this to another file e.g. saving the output
```


```{r Load spreadsheet with unique seafood species translations}
# Load master spreadsheet with species names and revised species names
MCS_translations <- vroom(file=paste(filepath,"SeafoodSpeciesCommonNames.csv", sep="/"))

# Check structure of the translation document
#str(MCS_translations)
```



### A) ii) Capture fisheries (MMO) UK vessels landing into the UK and abroad

Load production data on landings into the UK by UK and foreign vessels. 


```{r Specify the desired file path}
# Select file path with raw data
filepath <- paste(data_dir,"RawData/csv-files", sep = "") 
```


```{r Load data (2017 to 2020)}
# i) Landings into the UK by UK and foreign vessels: 2017 to 2021
# Read in .csv file
df_UKV_LandingUKAbroad_201721_data <- vroom(file=paste(filepath,"MMO_UKV_landingUKAbroad-201721.csv", sep="/"), na = c(".."))
 
# Check for missing Volumes. Right column summed to be total number of rows in the data set
md.pattern(df_UKV_LandingUKAbroad_201721_data, rotate.names = TRUE)

# Select desired variables and years (2017 to 2020)
df_UKV_LandingUKAbroad_201720_data <- df_UKV_LandingUKAbroad_201721_data[c("Species", "2017", "2018", "2019", "2020")]
```


```{r Transform data set (2017 to 2020)}
# Change structure of the data set 
df_UKV_LandingUKAbroad_201720_long <- melt(df_UKV_LandingUKAbroad_201720_data, id.vars = c("Species"), variable.name = "Year", value.name = "Volume")
```


```{r Load data (2012 to 2016)}
# ii) Landings into the UK by UK and foreign vessels: 2012 to 2016 
# Read in .csv file
df_UKV_LandingUKAbroad_201216_data <- vroom(file=paste(filepath,"MMO_UKV_landingUKAbroad-201216.csv", sep="/"), na = c(".."))

# Check for missing Volumes
md.pattern(df_UKV_LandingUKAbroad_201216_data, rotate.names = TRUE)

```


```{r Transform data set (2012 to 2016)}
# Change structure of the data set 
df_UKV_LandingUKAbroad_201216_long <- melt(df_UKV_LandingUKAbroad_201216_data, id.vars = c("Species"), variable.name = "Year", value.name = "Volume")
```


```{r Load data (2008 to 2011)}
# iii) Landings into the UK by UK and foreign vessels: 2008 to 2012 but select years 2009 to 2011. Always best to use the most recent data set.
# Read in .csv file
df_UKV_LandingUKAbroad_200913_data <- vroom(file=paste(filepath,"MMO_UKV_landingUKAbroad-200913.csv", sep="/"), na = c(".."))

# Check for missing Volumes
md.pattern(df_UKV_LandingUKAbroad_200913_data, rotate.names = TRUE)

# Select desired variables and years (2017 to 2019)
df_UKV_LandingUKAbroad_200911_data <- df_UKV_LandingUKAbroad_200913_data[c("Species", "2009", "2010","2011")]

```


```{r Transform data set (2009 to 2011)}
# Change structure of the data set 
df_UKV_LandingUKAbroad_200911_long <- melt(df_UKV_LandingUKAbroad_200911_data, id.vars = c("Species"), variable.name = "Year", value.name = "Volume")

```


```{r Combine data sets: i) 2008 to 2011; ii) 2012 to 2016; and iii) 2017 to 2019)}
# Combine capture production data i) 2009 to 2011; ii) 2012 to 2016; and iii) 2017 to 2020)
df_UKV_LandingUKAbroad_200920_long <- rbind(df_UKV_LandingUKAbroad_201720_long,
                              df_UKV_LandingUKAbroad_201216_long, 
                              df_UKV_LandingUKAbroad_200911_long)
```



Capture fisheries data are now all in one data frame. Data frame is composed of four columns: SpeciesType, Species, Year and Volume. 

We need to ensure species are spelled the same across all data sets, otherwise data sets cannot be merged, for instance "Blue whiting" v.s. "Blue Whiting or "Crab" v.s. "Crabs". We created an excel document listing the unique spellings of all species in all data sets used. In the cleaning process, the data sheets are joined to the "translation" document by the unique species name. In the final cleaned version, the correct species name column is selected. The translation document also provides various species type classifications as per published peer-reviewed papers.


```{r Join data frame with translation document}
# Join data frame with translation document- ensures consistent spelling across data sets
# Duplicate species name column (allows us to check species have been renames correctly)
df_UKV_LandingUKAbroad_200920_long$SpeciesRevised <- df_UKV_LandingUKAbroad_200920_long$Species

# Join data frame with MCS data frame by species name
df_UKV_LandingUKAbroad_200920_long <- df_UKV_LandingUKAbroad_200920_long %>% inner_join(., MCS_translations, by = c('SpeciesRevised' = 'UniqueSpecies'))
# MSC (DO NOT TOUCH) column is the unique spelling of the species in all data sets.

# Check column names of the updated data frame
colnames(df_UKV_LandingUKAbroad_200920_long)

```


Landings provided as whole weight. Need to apply conversion factors estimate edible portion of the seafood. 

```{r Apply conversion factors to calculate edible portion}
# Identify unique conversion factors in the data set
unique(df_UKV_LandingUKAbroad_200920_long$CFValue) # no processed data here so no 1 Volumes

# Create new column and apply conversion factors Volumes
df_UKV_LandingUKAbroad_200920_long$apply_CF <- df_UKV_LandingUKAbroad_200920_long$Volume * df_UKV_LandingUKAbroad_200920_long$CFValue

```


```{r Convert from 1000 tonnes to grams}
# Convert raw Volume from 000's tonnes to grams 
df_UKV_LandingUKAbroad_200920_long$Volume <- (df_UKV_LandingUKAbroad_200920_long$apply_CF * 1000) * 1000000
```


```{r Add additional variables to data set}
# Create "cleaned" data frame, defining units and data source 
df_UKV_LandingUKAbroad_200920_long$Units <- "Grams"
df_UKV_LandingUKAbroad_200920_long$DataSupplier <- "MMO"
df_UKV_LandingUKAbroad_200920_long$DataSet <- "Sea Fisheries Statistics"
df_UKV_LandingUKAbroad_200920_long$Commodity <- "Production"
df_UKV_LandingUKAbroad_200920_long$Flag <- "ConV1"
df_UKV_LandingUKAbroad_200920_long$TemporalResolution <- "CalendarYear"
df_UKV_LandingUKAbroad_200920_long$FlagDescription <- "CF applied to all or part of the data"

```


```{r Select desired variables for the cleaned data set}
# Identify column names
colnames(df_UKV_LandingUKAbroad_200920_long)

# Select desired columns for cleaned data set-- 16  variables
df_UKV_LandingUKAbroad_200920_clean <- df_UKV_LandingUKAbroad_200920_long[c("DataSupplier", "DataSet", "Commodity", "CommonSpeciesName", "SpeciesType", "SACN", "Volume", "Units", "Year", "TemporalResolution", "Flag", "FlagDescription")]

# Check column names
colnames(df_UKV_LandingUKAbroad_200920_clean)

# Rename "SeafoodSpecies" column 
#df_UKV_LandingUKAbroad_200918_cleaned <- rename(df_UKV_LandingUKAbroad_200918_clean, c("SeafoodSpecies" = "Species"))

df_UKV_LandingUKAbroad_200920_cleaned <- df_UKV_LandingUKAbroad_200920_clean %>% rename(Species = CommonSpeciesName)

```


```{r Save cleaned data}
# Change file path to "ProcessedData". Can also use setwd()
filepath <- paste(data_dir,"ProcessedData/", sep = "")

# Stating row.names = FALSE removes the initial column of the new .csv file where the rows are numbered
write.csv(df_UKV_LandingUKAbroad_200920_cleaned, file = paste(filepath,"UKVessLandingUKAbroad_MMO_PreliminaryCleaned.csv", sep = ""), row.names = FALSE)

```




### A) iii) Capture fisheries (MMO) UK vessels landing in to the UK only

Here we load landings data on landings into the UK by UK vessels. Multiple data sets are loaded to accommodate for the time period.


```{r Specify the desired file path}
# Select file path with raw data
filepath <- paste(data_dir,"RawData/csv-files", sep = "") 
```



```{r Load data (2017 to 2019)}
# i) Landings into the UK by UK and foreign vessels: 2017 to 2021
# Read in .csv file
df_UKV_LandingUKonly_201721_data <- vroom(file=paste(filepath,"MMO_UKVLandingUKOnly-201721.csv", sep="/"), na = c(".."))
 
# Check for missing Volumes. Right column summed to be total number of rows in the data set
md.pattern(df_UKV_LandingUKonly_201721_data, rotate.names = TRUE)

# Select desired variables and years (2017 to 2020)
df_UKV_LandingUKonly_201720_data <- df_UKV_LandingUKonly_201721_data[c("Species", "2017", "2018", "2019", "2020")]

```


```{r Transform data set (2017 to 2020)}
# Change structure of the data set 
df_UKV_LandingUKonly_201720_long <- melt(df_UKV_LandingUKonly_201720_data, id.vars = c("Species"), variable.name = "Year", value.name = "Volume")
```


```{r Load data (2012 to 2016)}
# ii) Landings into the UK by UK and foreign vessels: 2012 to 2016 
# Read in .csv file
df_UKV_LandingUKonly_201216_data <- vroom(file=paste(filepath,"MMO_UKVLandingUKOnly-201216.csv", sep="/"), na = c(".."))

# Check for missing Volumes
md.pattern(df_UKV_LandingUKonly_201216_data, rotate.names = TRUE)

# Select desired variables and years (2012 to 2016)
df_UKV_LandingUKonly_201216_data <- df_UKV_LandingUKonly_201216_data[c("Species", "2012", "2013", "2014", "2015", "2016")]

```


```{r Transform data set (2012 to 2016)}
# Change structure of the data set 
df_UKV_LandingUKonly_201216_long <- melt(df_UKV_LandingUKonly_201216_data, id.vars = c("Species"), variable.name = "Year", value.name = "Volume")
```


```{r Load data (2008 to 2011)}
# iii) Landings into the UK by UK and foreign vessels: 2008 to 2012 but select years 2009 to 2011. Always best to use the most recent data set.
# Read in .csv file
df_UKV_LandingUKonly_200812_data <- vroom(file=paste(filepath,"MMO_UKVLandingUKOnly-200812.csv", sep="/"), na = c(".."))

# Check for missing Volumes
md.pattern(df_UKV_LandingUKonly_200812_data, rotate.names = TRUE)

# Select desired variables and years (2017 to 2019)
df_UKV_LandingUKonly_200911_data <- df_UKV_LandingUKonly_200812_data[c("Species", "2009", "2010","2011")]

```


```{r Transform data set (2009 to 2011)}
# Change structure of the data set 
df_UKV_LandingUKonly_200911_long <- melt(df_UKV_LandingUKonly_200911_data, id.vars = c("Species"), variable.name = "Year", value.name = "Volume")

```


```{r Combine data sets: i) 2008 to 2011; ii) 2012 to 2016; and iii) 2017 to 2020)}
# Combine capture production data i) 2009 to 2011; ii) 2012 to 2016; and iii) 2017 to 2020)
df_UKV_LandingUKonly_200920_long <- rbind(df_UKV_LandingUKonly_201720_long,
                              df_UKV_LandingUKonly_201216_long, 
                              df_UKV_LandingUKonly_200911_long)
```



Capture fisheries data are now all in one data frame. Data frame is composed of four columns: SpeciesType, Species, Year and Volume. 

We need to ensure species are spelled the same across all data sets, otherwise data sets cannot be merged, for instance "Blue whiting" v.s. "Blue Whiting or "Crab" v.s. "Crabs". We created an excel document listing the unique spellings of all species in all data sets used. In the cleaning process, the data sheets are joined to the "translation" document by the unique species name. In the final cleaned version, the correct species name column is selected. The translation document also provides various species type classifications as per published peer-reviewed papers.


```{r Join data frame with translation document}
# Join data frame with translation document- ensures consistent spelling across data sets
# Duplicate species name column (allows us to check species have been renames correctly)
df_UKV_LandingUKonly_200920_long$SpeciesRevised <- df_UKV_LandingUKonly_200920_long$Species

# Subset the data frame so just MMO data only
MMO_MCS_translations <- subset(MCS_translations, DataSupplier == "MMO")

# Join data frame with MCS data frame by species name
df_UKV_LandingUKonly_200920_long <- df_UKV_LandingUKonly_200920_long %>% inner_join(., MMO_MCS_translations, by = c('SpeciesRevised' = 'UniqueSpecies'))

# Check column names of the updated data frame
colnames(df_UKV_LandingUKonly_200920_long)

```

Landings provided as whole weight. Need to apply conversion factors estimate edible portion of the seafood. 

```{r Apply conversion factors to calculate edible portion}
# Identify unique conversion factors in the data set
unique(df_UKV_LandingUKonly_200920_long$CFValue) # no processed data here so no 1 Volumes

# Create new column and apply conversion factors Volumes
df_UKV_LandingUKonly_200920_long$apply_CF <- df_UKV_LandingUKonly_200920_long$Volume * df_UKV_LandingUKonly_200920_long$CFValue

```


```{r Convert from 1000 tonnes to grams}
# Convert raw Volume from 000's tonnes to grams 
df_UKV_LandingUKonly_200920_long$Volume <- (df_UKV_LandingUKonly_200920_long$apply_CF * 1000) * 1000000
```


```{r Add additional variables to data set}
# Create "cleaned" data frame, defining units and data source 
df_UKV_LandingUKonly_200920_long$Units <- "Grams"
df_UKV_LandingUKonly_200920_long$DataSupplier <- "MMO"
df_UKV_LandingUKonly_200920_long$DataSet <- "Sea Fisheries Statistics"
df_UKV_LandingUKonly_200920_long$Commodity <- "Production"
df_UKV_LandingUKonly_200920_long$Flag <- "ConV1"
df_UKV_LandingUKonly_200920_long$TemporalResolution <- "CalendarYear"
df_UKV_LandingUKonly_200920_long$FlagDescription <- "CF applied to all or part of the data"

```


```{r Select desired variables for the cleaned data set}
# Identify column names
colnames(df_UKV_LandingUKonly_200920_long)

# Select desired columns for cleaned data set-- 16  variables
df_UKV_LandingUKonly_200920 <- df_UKV_LandingUKonly_200920_long[c("DataSupplier", "DataSet", "Commodity", "CommonSpeciesName", "SpeciesType", "SACN", "Volume", "Units", "Year", "TemporalResolution", "Flag", "FlagDescription")]

# Rename "SeafoodSpecies" column 
#df_UKV_LandingUKAbroad_200918_cleaned <- rename(df_UKV_LandingUKAbroad_200918_clean, c("SeafoodSpecies" = "Species"))

df_UKV_LandingUKonly_200920_cleaned <- df_UKV_LandingUKonly_200920 %>% rename(Species = CommonSpeciesName)

```


```{r Check dataset}
# Check for duplicates
dulpicates_MMO <- df_UKV_LandingUKonly_200920_cleaned %>% select(DataSupplier, DataSet, Commodity, Species, Year, Volume, Units) %>% group_by(DataSupplier, DataSet, Commodity, Species, Year, Volume, Units) %>% 
  filter(n()>1) %>% mutate(n = n())

# Check data transforms from long and thin to short and wide
shape_MMO <- df_UKV_LandingUKonly_200920_cleaned %>% 
  select(DataSupplier, DataSet, Commodity, Species, SACN, Year, Volume, Units) %>% 
  filter(Commodity %in% c("Production")) %>% 
  pivot_wider(names_from = Commodity, values_from = Volume)

```


```{r Save cleaned data}
# Change file path to "ProcessedData". Can also use setwd()
filepath <- paste(data_dir,"ProcessedData/", sep = "")

# Stating row.names = FALSE removes the initial column of the new .csv file where the rows are numbered
write.csv(df_UKV_LandingUKonly_200920_cleaned, file = paste(filepath,"OnlyUKVessels_LandingUK_MMO_PreliminaryCleaned.csv", sep = ""), row.names = FALSE)

```

