---
title: "Cleaning_Purchase-data_DEFRA"
author: "AL"
date: "`r Sys.Date()`"
output: html_document
---


```{r setup, include=FALSE}
# sets options for entire document in the first chunk
knitr::opts_chunk$set(echo = TRUE)
```

## Preparation

```{r Load libraries, include=FALSE}
# Load packages
library(tidyr) # for tidying data
library(dplyr) # base R alternative but neater
library(vroom) # for loading and transforming data
library(tidyverse) # data exploration 
library(reshape2) # for melting data frames i.e. short and wide to long and thin
library(data.table) # for fread()-function
library(mice) # md.pattern to show missing data
library(stringr) # used to replace matched patterns in a string
library(haven) #Read SAV files
library(readxl)
library(readr) # CSV file I/O, e.g. the read_csv function
library(sjlabelled)
library(compareDF)#Check missing values across datasets
library(chron)

```


```{r Load data, echo=TRUE}
# Loading file path to the data into a Volume 
source("Data_filepath.R")# Data_filepath.R is listed in .gitignore-file. So, you will need to create that file yourself and provide your respected filepath using "data_dir <- [enter your here]"
```


```{r Specify the desired file path}
filepath <- paste(data_dir,"Methods/SpeciesTypeClassification", sep = "") 
# note we are able to change this to another file e.g. saving the output
```


```{r Load spreadsheet with unique seafood species translations}
# Load master spreadsheet with species names and revised species names
MCS_translations <- vroom(file=paste(filepath,"SeafoodSpeciesCommonNames.csv", sep="/"))

# Check structure of the translation document
#str(MCS_translations)
```


```{r Specify the desired file path}
filepath <- paste(data_dir,"ProcessedData", sep = "") 
# note we are able to change this to another file e.g. saving the output
```


```{r Load UK population data}
# Load master spreadsheet with species names and revised species names
#MCS_translations <- vroom(file=paste(filepath,"SeafoodSpeciesCommonNames.csv", sep="/"))
df_UK_population_200920 <- vroom(file=paste(filepath,"UKPopulation_ONS_PreliminaryCleaned.csv", sep="/"))

# Check structure of the translation document
str(df_UK_population_200920)
```



## Purchases (DEFRA family food) UK seafood purchases

Load  UK seafood purchasing data. Purchasing data between 2016 and 2020 is composed of purchases at a household and eaten out level. The two data sets need to be separated. 


```{r Specify the desired file path}
# Select file path with raw data
filepath <- paste(data_dir,"RawData/csv-files", sep = "") 
```


```{r Load data (2016 to 2020)} 
# *Error in raw data in 2015, so data from 2016 onwards used

# read DEFRA purchasing (household and eating out) data (.csv)
df_DEFRA_purchase_201620_data <- vroom(file=paste(filepath,"DEFRA_UK_purchases-201620.csv", sep="/"))

# Check for missing Volumes
md.pattern(df_DEFRA_purchase_201620_data)

# State full calendar year date
df_DEFRA_purchase_201620_data$Year[df_DEFRA_purchase_201620_data$Year==15] <- 2015
df_DEFRA_purchase_201620_data$Year[df_DEFRA_purchase_201620_data$Year==16] <- 2016
df_DEFRA_purchase_201620_data$Year[df_DEFRA_purchase_201620_data$Year==17] <- 2017
df_DEFRA_purchase_201620_data$Year[df_DEFRA_purchase_201620_data$Year==18] <- 2018
df_DEFRA_purchase_201620_data$Year[df_DEFRA_purchase_201620_data$Year==19] <- 2019
df_DEFRA_purchase_201620_data$Year[df_DEFRA_purchase_201620_data$Year==20] <- 2020
unique(df_DEFRA_purchase_201620_data$Year)

# Subset years 2016 to 2020
df_DEFRA_purchase_201620_data <- subset(df_DEFRA_purchase_201620_data,  Year > 2015)
unique(df_DEFRA_purchase_201620_data$Year)
```


Need to join seafood purchasing data with the seafood translation document before subsetting- can then subset data by data set name.

```{r Join data frame with translation document}
# Duplicate species name column (allows us to check species have been renames correctly)
df_DEFRA_purchase_201620_data$SpeciesRevised <- df_DEFRA_purchase_201620_data$codeDESCRIPTION

# Join data frame with MCS data frame by species name
df_DEFRA_purchase_201620_data <- df_DEFRA_purchase_201620_data %>% inner_join(., MCS_translations, by = c('SpeciesRevised' = 'UniqueSpecies'))
# MSC (DO NOT TOUCH) column is the unique spelling of the species in all data sets.

# Check column names of the updated data frame
colnames(df_DEFRA_purchase_201620_data)

```


```{r Subset purchasing data} 
# Subset purchasing data for household and eating out purposes based on data set name
df_DEFRA_household_201620 <- subset(df_DEFRA_purchase_201620_data, DataSupplier == "DEFRA_household") 
df_DEFRA_eatenOut_201620 <- subset(df_DEFRA_purchase_201620_data, DataSupplier == "DEFRA_eatenOut") 

```


Household and eaten-out purchase data are now separated.


### ii) Household data (2016 to 2020)

```{r Transform data (household 2016 to 2020)}
# Select only years 2016 to 2020
df_DEFRA_household_201620_short <- subset(df_DEFRA_purchase_201620_data, 
                           select = c("Year", "codeDESCRIPTION", "estimate"))
unique(df_DEFRA_household_201620_short$codeDESCRIPTION)

df_DEFRA_household_201620_short$Units <- "g"

```


### iii) Household data (2009 to 2015)

Load household purchases between 2009 and 2015. Note that the structures are different between the data sets. 

```{r Load data (household 2009 to 2015)}
# read DEFRA household data set (.csv)
df_DEFRA_household_200915_data <- vroom(file=paste(filepath,"DEFRA_UK_household-200915.csv", sep="/"))

# Select only years 2009 to 2015
df_DEFRA_household_200915 <- subset(df_DEFRA_household_200915_data, 
                           select = c("MinorFoodCode", "Units", "2009", "2010", "2011", 
                                      "2012", "2013", "2014","2015"))
# Check for missing Volumes
#md.pattern(df_DEFRA_household_200915, rotate.names = TRUE)

```


```{r Transform data set (2009 to 2015)}
# Change structure of the data set 
df_DEFRA_household_200915_long <- melt(df_DEFRA_household_200915, id.vars = c("MinorFoodCode", "Units"), variable.name ="Year", value.name = "Volume")
```


Now combine the data sets. 

```{r Combine DEFRA household purchaing data i) 2019 to 2015 ii) 2016 to 2020}
# Check data frame structure  
colnames(df_DEFRA_household_201620_short)
colnames(df_DEFRA_household_200915_long)

# Rename variables to allow data frames to merge
# Rename codeDescription and estimate column
df_DEFRA_household_201620_to_join <- rename(df_DEFRA_household_201620_short, 
                                            FoodProduct = codeDESCRIPTION, Volume = estimate)

# Rename minor food code column
df_DEFRA_household_201915_to_join <- rename(df_DEFRA_household_200915_long, FoodProduct = MinorFoodCode)

# Ensure "Year" column is a numerical Volume. Convert to character first, then numeric
df_DEFRA_household_201915_to_join$Year <- as.numeric(as.character(df_DEFRA_household_201915_to_join$Year))
df_DEFRA_household_201620_to_join$Year <- as.numeric(as.character(df_DEFRA_household_201620_to_join$Year))

# Combine data frames
df_DEFRA_household_200920 <- rbind(df_DEFRA_household_201915_to_join, df_DEFRA_household_201620_to_join)
```


```{r Delete Later: volumes purchased per household}
str(df_DEFRA_household_200920)
df_DEFRA_household_200920_short <- aggregate(Volume ~ Year, data = df_DEFRA_household_200920, sum)
plot(Volume ~ Year, data = df_DEFRA_household_200920_short)

```


```{r Join data frame with translation document}
# Duplicate species name column (allows us to check species have been renames correctly)
df_DEFRA_household_200920$SpeciesRevised <- df_DEFRA_household_200920$FoodProduct

# Join data frame with MCS data frame by species name
df_DEFRA_household_200920 <- df_DEFRA_household_200920 %>% inner_join(., MCS_translations, by = c('SpeciesRevised' = 'UniqueSpecies'))
# MSC (DO NOT TOUCH) column is the unique spelling of the species in all data sets.

# Observations in the data frame will be less as original table included "sum" of some species. 

# Check column names
colnames(df_DEFRA_household_200920)

```


```{r Seafood consumption (g/capita/week), by species}
# Sum seafood consumption per person per species
df_DEFRA_household_200920_agg <- aggregate(Volume ~ CommonSpeciesName + Year + SpeciesType + SACN, data =
                                                df_DEFRA_household_200920, sum)

```


```{r Convert g/capita/week to g/country/year}
# Merge household purchasing data with population data 
df_DEFRA_household_200920_aggYear <- merge(df_DEFRA_household_200920_agg, df_UK_population_200920, "Year")

# Multiply Volume data by 52 (convert grams from week to year)
df_DEFRA_household_200920_aggYear$PerYear <- df_DEFRA_household_200920_aggYear$Volume * 52

# Create new column multiplying by population size
df_DEFRA_household_200920_aggYear$totUK <- df_DEFRA_household_200920_aggYear$PerYear * df_DEFRA_household_200920_aggYear$UKPOP

str(df_DEFRA_household_200920_aggYear)

```


```{r Add additional variables to data set}
# Add additional variables
df_DEFRA_household_200920_aggYear$DataSupplier <- "DEFRA"
df_DEFRA_household_200920_aggYear$DataSet <- "Family Foods Household"
df_DEFRA_household_200920_aggYear$Commodity <- "Purchases"
df_DEFRA_household_200920_aggYear$Flag <- "ConV2"
df_DEFRA_household_200920_aggYear$TemporalResolution <- "CalendarYear"
df_DEFRA_household_200920_aggYear$Units <- "Grams"
df_DEFRA_household_200920_aggYear$FlagDescription <- "Unit converted from g/capita/week to g/country/year"

```


```{r Select desired variables for the data set}
# Identify columns in data frame
colnames(df_DEFRA_household_200920_aggYear)

# Select desired columns for cleaned data set
df_DEFRA_household_200920_cleaned <- df_DEFRA_household_200920_aggYear[c("DataSupplier", "DataSet", "Commodity", "CommonSpeciesName", "SpeciesType", "SACN", "totUK", "Units", "Year", "TemporalResolution", "Flag", "FlagDescription")]

# Rename some variables
df_DEFRA_household_200920_aggYear_cleaned <- rename(df_DEFRA_household_200920_cleaned, 
                                                    Species = CommonSpeciesName, Volume = totUK)

```


```{r Check dataset}
# Check for duplicates
duplicates_DEFRAHouse <- df_DEFRA_household_200920_aggYear_cleaned %>% select(DataSupplier, DataSet, Commodity, Species, Year, Volume, Units) %>% group_by(DataSupplier, DataSet, Commodity, Species, Year, Volume, Units) %>% 
  filter(n()>1) %>% mutate(n = n())

# Check data transforms from long and thin to short and wide
shape_DEFRAHouse <- df_DEFRA_household_200920_aggYear_cleaned %>% 
  select(DataSupplier, DataSet, Commodity, Species, SACN, Year, Volume, Units) %>% 
  filter(Commodity %in% c("Purchases")) %>% 
  pivot_wider(names_from = Commodity, values_from = Volume)

```


```{r Save cleaned data}
# Change file path to "ProcessedData"
filepath <- paste(data_dir,"ProcessedData/", sep="")
write.csv(df_DEFRA_household_200920_aggYear_cleaned, file =
            paste(filepath,"HouseholdPurchases_DEFRA_PreliminaryCleaned.csv", sep=""), row.names = FALSE)

```



### iii) Eating out (2016 to 2020)


```{r Transform data (household 2016 to 2020)}
# Select only years 2016 to 2020
df_DEFRA_eatenOut_201620_short <- subset(df_DEFRA_eatenOut_201620, 
                           select = c("Year", "codeDESCRIPTION", "estimate"))

# Allocate units
df_DEFRA_eatenOut_201620_short$Units <- "g/capita/week"

```


```{r Delete Later: purchases eaten out}
str(df_DEFRA_eatenOut_201620_short)
df_DEFRA_eatenOut_201620_short_shorter <- aggregate(estimate ~ Year, data = df_DEFRA_eatenOut_201620_short, sum)
plot(estimate ~ Year, data = df_DEFRA_eatenOut_201620_short_shorter)
```


### iv) Eating out (2009 to 2015)


```{r Specify the desired file path}
# Select file path with raw data
filepath <- paste(data_dir,"RawData/csv-files", sep = "") 
```


```{r Load data (2009 to 2015)}
# read DEFRA eaten out data set
df_DEFRA_eatenOut_200915_data <- vroom(file = paste(filepath,"DEFRA_UK_eatenOut-200915.csv", sep="/"))

# Check for missing Volumes
#md.pattern(df_DEFRA_eatenOut_200915_data, rotate.names = T)
# Missing Volumes detected but not within time frame of interest

# Select only years 2009 to 2015
df_DEFRA_eatenOut_200915 <- subset(df_DEFRA_eatenOut_200915_data, 
                           select = c("EO_MinorCode", "Units", "2009", "2010", "2011",
                                      "2012", "2013","2014", "2015"))

```


```{r Transform data set (2009 to 2015)}
# Change structure of the data set 
df_DEFRA_eatenOut_200915_long <- melt(df_DEFRA_eatenOut_200915, id.vars = c("EO_MinorCode", "Units"), variable.name = "Year", value.name = "Volume")

```



```{r Combine DEFRA household purchaing data i) 2019 to 2015 ii) 2016 to 2020}
# Check data frame structure  
colnames(df_DEFRA_eatenOut_200915_long)
colnames(df_DEFRA_eatenOut_201620_short)

# Rename variables to allow data frames to merge
# Rename codeDescription and estimate column
df_DEFRA_eatenOut_201620_to_join <- rename(df_DEFRA_eatenOut_201620_short, 
                                            FoodProduct = codeDESCRIPTION, Volume = estimate)

# Rename minor food code column
df_DEFRA_eatenOut_201915_to_join <- rename(df_DEFRA_eatenOut_200915_long, FoodProduct = EO_MinorCode)

# Ensure "Year" column is a numerical Volume. Convert to character first, then numeric
df_DEFRA_eatenOut_201915_to_join$Year <- as.numeric(as.character(df_DEFRA_eatenOut_201915_to_join$Year))
df_DEFRA_eatenOut_201620_to_join$Year <- as.numeric(as.character(df_DEFRA_eatenOut_201620_to_join$Year))

# Combine data frames
df_DEFRA_eatenOut_200920 <- rbind(df_DEFRA_eatenOut_201915_to_join, df_DEFRA_eatenOut_201620_to_join)

```




```{r Join data frame with translation document}
# Duplicate species name column (allows us to check species have been renames correctly)
df_DEFRA_eatenOut_200920$SpeciesRevised <- df_DEFRA_eatenOut_200920$FoodProduct

# Join data frame with MCS data frame by species name
df_DEFRA_eatenOut_200920 <- df_DEFRA_eatenOut_200920 %>% inner_join(., MCS_translations, by = c('SpeciesRevised' = 'UniqueSpecies'))
# MSC (DO NOT TOUCH) column is the unique spelling of the species in all data sets.

# Check column names of the updated data frame
colnames(df_DEFRA_eatenOut_200920)

```


```{r Seafood consumption (g/capita/week), by species}
# Sum seafood consumption per person per species
df_DEFRA_eatenOut_200920_agg <- aggregate(Volume ~ CommonSpeciesName + Year + SpeciesType + SACN, data =
                                                df_DEFRA_eatenOut_200920, sum)

```


```{r Convert g/capita/week to g/country/year}
# Merge household purchasing data with population data 
df_DEFRA_eatenOut_200920_aggYear <- merge(df_DEFRA_eatenOut_200920_agg, df_UK_population_200920, "Year")

# Multiply Volume data by 52 (convert grams from week to year)
df_DEFRA_eatenOut_200920_aggYear$PerYear <- df_DEFRA_eatenOut_200920_aggYear$Volume * 52

# Create new column multiplying by population size
df_DEFRA_eatenOut_200920_aggYear$totUK <- df_DEFRA_eatenOut_200920_aggYear$PerYear * df_DEFRA_eatenOut_200920_aggYear$UKPOP

str(df_DEFRA_eatenOut_200920_aggYear)

```


```{r Add additional variables to data set}
# Add additional variables
df_DEFRA_eatenOut_200920_aggYear$DataSupplier <- "DEFRA"
df_DEFRA_eatenOut_200920_aggYear$DataSet <- "Family Foods Eaten Out"
df_DEFRA_eatenOut_200920_aggYear$Commodity <- "Purchases"
df_DEFRA_eatenOut_200920_aggYear$Flag <- "ConV2"
df_DEFRA_eatenOut_200920_aggYear$TemporalResolution <- "CalendarYear"
df_DEFRA_eatenOut_200920_aggYear$Units <- "Grams"
df_DEFRA_eatenOut_200920_aggYear$FlagDescription <- "Unit converted from g/capita/week to g/country/year"

```


```{r Select desired variables for the data set}
# Identify columns in data frame
colnames(df_DEFRA_eatenOut_200920_aggYear)

# Select desired columns for cleaned data set
df_DEFRA_eatenOut_200920_cleaned <- df_DEFRA_eatenOut_200920_aggYear[c("DataSupplier", "DataSet", "Commodity", "CommonSpeciesName", "SpeciesType", "SACN", "totUK", "Units", "Year", "TemporalResolution", "Flag", "FlagDescription")]

# Rename some variables
df_DEFRA_eatenOut_200920_cleaned <- rename(df_DEFRA_eatenOut_200920_cleaned, 
                                                    Species = CommonSpeciesName, Volume = totUK)

colnames(df_DEFRA_eatenOut_200920_cleaned)

```



```{r Check dataset}
# Check for duplicates
duplicates_DEFRAOut <- df_DEFRA_eatenOut_200920_cleaned %>% select(DataSupplier, DataSet, Commodity, Species, Year, Volume, Units) %>% group_by(DataSupplier, DataSet, Commodity, Species, Year, Volume, Units) %>% 
  filter(n()>1) %>% mutate(n = n())

# Check data transforms from long and thin to short and wide
shape_DEFRAOut <- df_DEFRA_eatenOut_200920_cleaned %>% 
  select(DataSupplier, DataSet, Commodity, Species, SACN, Year, Volume, Units) %>% 
  filter(Commodity %in% c("Purchases")) %>% 
  pivot_wider(names_from = Commodity, values_from = Volume)

```



```{r Save cleaned data}
# Change file path to "ProcessedData"
filepath <- paste(data_dir,"ProcessedData/", sep="")
write.csv(df_DEFRA_eatenOut_200920_cleaned, file =
            paste(filepath,"EatenOutPurchases_DEFRA_PreliminaryCleaned.csv", sep=""), row.names = FALSE)

```


