---
title: "Cleaning_Consumption-data_NDNS"
author: "AL"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
# sets options for entire document in the first chunk
knitr::opts_chunk$set(echo = TRUE)
```

## Preparation

```{r Load libraries, include=FALSE}
# Load packages
library(tidyr) # for tidying data
library(dplyr) # base R alternative but neater
library(vroom) # for loading and transforming data
library(tidyverse) # data exploration 
library(reshape2) # for melting data frames i.e. short and wide to long and thin
library(data.table) # for fread()-function
library(mice) # md.pattern to show missing data
library(stringr) # used to replace matched patterns in a string
library(haven) #Read SAV files
library(readxl)
library(readr) # CSV file I/O, e.g. the read_csv function
library(sjlabelled)
library(compareDF)#Check missing values across datasets
library(chron)

```


```{r Load data, echo=TRUE}
# Loading file path to the data into a Volume 
source("Data_filepath.R")# Data_filepath.R is listed in .gitignore-file. So, you will need to create that file yourself and provide your respected filepath using "data_dir <- [enter your here]"
```


```{r Specify the desired file path}
filepath <- paste(data_dir,"Methods/SpeciesTypeClassification", sep = "") 
# note we are able to change this to another file e.g. saving the output
```


```{r Load spreadsheet with unique seafood species translations}
# Load master spreadsheet with species names and revised species names
MCS_translations <- vroom(file=paste(filepath,"SeafoodSpeciesCommonNames.csv", sep="/"))

# Check structure of the translation document
#str(MCS_translations)
```



```{r Specify the desired file path}
filepath <- paste(data_dir,"ProcessedData", sep = "") 
# note we are able to change this to another file e.g. saving the output
```


```{r Load UK population data}
# Load master spreadsheet with species names and revised species names
#MCS_translations <- vroom(file=paste(filepath,"SeafoodSpeciesCommonNames.csv", sep="/"))
df_UK_population_200920 <- vroom(file=paste(filepath,"UKPopulation_ONS_PreliminaryCleaned.csv", sep="/"))

# Check structure of the translation document
str(df_UK_population_200920)
```



## Consumption (NDNS) UK seafood cosnumption to food level

UK fish consumption data was sourced from the National Diet and Nutrition Survey (NDNS). Food diaries show the amount of seafood and often species consumed in the UK. It has greater granularity compared to other data sources (consumption identified to species level).

UK seafood consumption data is collected every financial year. Although food diary entry dates are provided, yearly consumption per calendar year cannot be calculated due to the nature of the data collection. Therefore, over a 10-year period, there will be five seafood consumption estimates. NDNS data is freely available from 2008 to 2019 (i.e. reporting years 1 to 11). Data, to food level consumption can be downloaded as txt files from the UK data service. Convert all tab files to .csv before loading into R. 


```{r Specify the desired file path}
# Select file path with raw data
filepath <- paste(data_dir,"RawData/csv-files", sep = "") 
```


Download from SPSS.

```{r Load data (2008 to 2019)}
# Load NDNS yrs 1 - 4
df_NDNS_foodlevel_yr1_4_data <- 
  read_sav(file= paste(filepath,"ndns_rp_yr1-4a_foodleveldietarydata_uk_v2.sav", sep = "/"))

# Load NDNS yrs 5 - 6 
df_NDNS_foodlevel_yr5_6_data <-
  read_sav(file = paste(filepath,"ndns_rp_yr5-6a_foodleveldietarydata_v2.sav", sep = "/"))

# Load NDNS yrs 7 - 8 
df_NDNS_foodlevel_yr7_8_data <-
  read_sav(file = paste(filepath,"ndns_rp_yr7-8a_foodleveldietarydata.sav", sep = "/"))

# Load NDNS yrs 9
df_NDNS_foodlevel_yr9_data <- 
  read_sav(file = paste(filepath,"ndns_rp_yr9a_foodleveldietarydata_uk.sav", sep = "/"))

# Load NDNS yrs 10
df_NDNS_foodlevel_yr10_data <- 
  read_sav(file = paste(filepath,"ndns_rp_yr10a_foodleveldietarydata_uk.sav", sep = "/"))

# Load NDNS yrs 11
df_NDNS_foodlevel_yr11_data <- 
  read_sav(file = paste(filepath,"ndns_rp_yr11a_foodleveldietarydata_uk.sav", sep = "/"))

# Load NDNS yrs 12
df_NDNS_foodlevel_yr12_data <- 
  read_sav(file = paste(filepath,"ndns_rp_19_foodleveldietarydata.sav", sep = "/"))


```


```{r Clean and combine data sets (Year 1 - 11) }
# Clean data 1- 11 NDNS
df_NDNS_foodlevel_yr1_4_short <- subset(df_NDNS_foodlevel_yr1_4_data, 
                                       select = -c(diarymth,WhoWithOther,WhereOther))

df_NDNS_foodlevel_yr5_6_short <- subset(df_NDNS_foodlevel_yr5_6_data, 
                                        select = -c(diarymth,WhoWithOther,WhereOther))

df_NDNS_foodlevel_yr7_8_short <- subset(df_NDNS_foodlevel_yr7_8_data, 
                                        select = -c(CoreBoost,diarymth))

df_NDNS_foodlevel_yr9_rename <- df_NDNS_foodlevel_yr9_data %>% rename(Age = AgeR)
df_NDNS_foodlevel_yr9_short <- subset(df_NDNS_foodlevel_yr9_rename, 
                                      select = -c(CoreBoost,DiaryDate,RecipeName))

df_NDNS_foodlevel_yr10_rename <- df_NDNS_foodlevel_yr10_data %>% rename(Age = AgeR)
df_NDNS_foodlevel_yr10_short <- subset(df_NDNS_foodlevel_yr10_rename, 
                                       select = -c(CoreBoost,DiaryDate,RecipeName))

df_NDNS_foodlevel_yr11_rename <- df_NDNS_foodlevel_yr11_data %>% rename(Age = AgeR)
df_NDNS_foodlevel_yr11_short <- subset(df_NDNS_foodlevel_yr11_rename, 
                                       select = -c(CoreBoost,DiaryDate,RecipeName))

```


```{r Join data frames}
# Join data frames
df_NDNS_UK_Yr1_11_data <- rbind(df_NDNS_foodlevel_yr1_4_short,
                            df_NDNS_foodlevel_yr5_6_short,
                            df_NDNS_foodlevel_yr7_8_short,
                            df_NDNS_foodlevel_yr9_short,
                            df_NDNS_foodlevel_yr10_short,
                            df_NDNS_foodlevel_yr11_short)

# Check for missing Volumes
#md.pattern(df_NDNS_UK_Yr1_11_data)
```


Downloaded NDNS data for all food groups. Need to select just those rows related to seafood consumption. Entries numbered 33, 34 and 35 "Main food group code" column related to seafood consumption.


```{r Subset seafood products from consumption data (Year 1 - 11)}
# Identify unique main food groups
unique(df_NDNS_UK_Yr1_11_data$MainFoodGroupCode)

# Subset seafood commodities. Seafood commodities identified as Volumes: 33, 34 and 35 in "main food group code" column
df_NDNS_seafood_consumption_Yr1_11 <- subset(df_NDNS_UK_Yr1_11_data, MainFoodGroupCode == 33 | df_NDNS_UK_Yr1_11_data$MainFoodGroupCode == 34| df_NDNS_UK_Yr1_11_data$MainFoodGroupCode == 35)

```


Calculate total grams of seafood consumed. Seafood intakes are spread across 4 columns. 

```{r Calculate total grams of seafood consummed (Year 1 - 11)}
# Calculate total seafood consumption. Sum columns"WhiteFishg","OilyFishg", "CannedTunag" and "Shellfishg"
df_NDNS_seafood_consumption_Yr1_11$totfishperprod <- df_NDNS_seafood_consumption_Yr1_11$WhiteFishg + df_NDNS_seafood_consumption_Yr1_11$OilyFishg + df_NDNS_seafood_consumption_Yr1_11$CannedTunag + df_NDNS_seafood_consumption_Yr1_11$Shellfishg
```


```{r Subset seafood products from consumption data (Year 12)}
# Identify unique food numbers
df_fishFoodnumber <- unique(df_NDNS_seafood_consumption_Yr1_11$FoodNumber)

# Subset fish products, using "food number" variable, from NDNS year 12 dataset
df_NDNS_sub_Yr12 <- df_NDNS_foodlevel_yr12_data[df_NDNS_foodlevel_yr12_data$FoodNumber %in% df_fishFoodnumber,]
```


```{r Calculate total grams of seafood consummed (Year 12)}
# Calcualte total grams of seafood consummed
df_NDNS_sub_Yr12$totfishperprod <- df_NDNS_sub_Yr12$WhiteFishg + df_NDNS_sub_Yr12$OilyFishg + df_NDNS_sub_Yr12$CannedTunag + df_NDNS_sub_Yr12$Shellfishg


#filepath <- paste(data_dir,"ProcessedData/", sep = "")
#write.csv(uniqueFishProductNames, file = paste(filepath,"NDNS_UniqueFishProd.csv", sep = ""), row.names = FALSE)

```


```{r Join data frame with translation document (Year 1-11)}
# Duplicate species name column (allows us to check species have been renames correctly)
df_NDNS_seafood_consumption_Yr1_11$SpeciesRevised <- df_NDNS_seafood_consumption_Yr1_11$FoodName

# Join data frame with MCS data frame by species name
df_NDNS_seafood_consumption_Yr1_11 <- df_NDNS_seafood_consumption_Yr1_11 %>% 
  inner_join(., MCS_translations, by = c('SpeciesRevised' = 'UniqueSpecies'))

# MSC (DO NOT TOUCH)/ Species is the unique spelling of the species in all data sets.
```


```{r Join data frame with translation document (Year 12)}
# NDNS year 12 does not provide species name so use food description.
df_NDNS_sub_Yr12$SpeciesRevised <- df_NDNS_sub_Yr12$FoodDescription

# Join data frame with MCS data frame by species name
df_NDNS_sub_Yr12 <- df_NDNS_sub_Yr12 %>% inner_join(., MCS_translations, by = c('SpeciesRevised' = 'UniqueSpecies')) # "Species" column is the unique spelling of the species in all data sets.

```


Now all species are in a separate column. Some species are duplicated per year i.e. salmon fillet/ salmon pate/ smoked salmon now all called salmon so need to aggregate (add) salmon consumption each reported year. So aggregate seafood consumption data, per species, per year. 


```{r Calculate daily fish intake per reported seafood consumer (Year 1 - 11)}
# Divide seafood consumption by diary days entered (total diary days completed changes per person)
df_NDNS_seafood_consumption_Yr1_11$gDay <- df_NDNS_seafood_consumption_Yr1_11$totfishperprod/ df_NDNS_seafood_consumption_Yr1_11$DiaryDaysCompleted
```


```{r Calculate daily fish intake per reported seafood consumer (Year 12)}
# In the new NDNS data DiaryDays completed is registered as RecallNo so summarise by seriali and get the max(RecallNo)- same as DiaryDaysCompleted. However, in year 12, not all the included individuals have 3 or 4 days unlike previous years.

DiaryDaysCompleted <- df_NDNS_sub_Yr12 %>% 
  group_by(seriali) %>%
  summarise(DiaryDaysCompleted = max(RecallNo)) 

# Merge DiaryDaysCompleted with the original data 
df_NDNS_sub_Yr12 <- merge(df_NDNS_sub_Yr12, DiaryDaysCompleted, by = "seriali")

# Now you can divide
df_NDNS_sub_Yr12$gDay <- df_NDNS_sub_Yr12$totfishperprod/ df_NDNS_sub_Yr12$DiaryDaysCompleted

```


```{r Sum daily seafood consumption per by species (both)}
# Sum seafood consumption per species per year- some respondents consumed fish more than once over the four days

#df_NDNS_seafood_UKReportedIntake_Yr1_11 <- aggregate(gDay ~ CommonSpeciesName + SurveyYear + SpeciesType + SACN, data = df_NDNS_seafood_consumption_Yr1_11, sum) # inclues species type

df_NDNS_seafood_UKReportedIntake_Yr1_11 <- aggregate(gDay ~ CommonSpeciesName + SurveyYear + SACN, data = df_NDNS_seafood_consumption_Yr1_11, sum)

df_NDNS_seafood_UKReportedIntake_Yr12 <- aggregate(gDay ~ CommonSpeciesName
                                                     + SACN, data = df_NDNS_sub_Yr12, sum)

```


```{r Calucaulte total number of NDNS participants per year (Year 1- 11)}
# Estimate total number of NDNS participants per year
respondersPerYear <- df_NDNS_UK_Yr1_11_data %>% 
  group_by(SurveyYear) %>%
    summarise(count = n_distinct(seriali)) 

# Merge NDNS participants with consumption data 
df_NDNS_seafoodReported_Yr1_11 <- merge(df_NDNS_seafood_UKReportedIntake_Yr1_11, respondersPerYear, by = "SurveyYear")
```


```{r Calucaulte daily and weekly UK seafood consumption (Year 1 - 11)}
# Divide seafood consumption by NDNS respondents 
df_NDNS_seafoodReported_Yr1_11$avgUKdailyIntake <- df_NDNS_seafoodReported_Yr1_11$gDay / df_NDNS_seafoodReported_Yr1_11$count


# Calculate weekly UK seafood consumption 
df_NDNS_seafoodReported_Yr1_11$avgUKWeeklyIntake <- df_NDNS_seafoodReported_Yr1_11$avgUKdailyIntake * 7  
```


```{r Calucaulte total number of NDNS participants per year (Year 12)}
# Add a column indicating it is year 12 to year 12
df_NDNS_seafood_UKReportedIntake_Yr12$SurveyYear <- "NDNS Year 12"
# Count serialis in year 12
respondersYear12 <- df_NDNS_foodlevel_yr12_data %>% 
  summarise(count = n_distinct(seriali)) 
# Estimate total number of NDNS participants per year
df_NDNS_seafood_UKReportedIntake_Yr12$count = respondersYear12$count
```


```{r Calucaulte daily and weekly UK seafood consumption (Year 12)}
# Divide seafood consumption by NDNS respondents 
df_NDNS_seafood_UKReportedIntake_Yr12$avgUKdailyIntake <- df_NDNS_seafood_UKReportedIntake_Yr12$gDay / df_NDNS_seafood_UKReportedIntake_Yr12$count
# Calculate weekly UK seafood consumption 
df_NDNS_seafood_UKReportedIntake_Yr12$avgUKWeeklyIntake <- df_NDNS_seafood_UKReportedIntake_Yr12$avgUKdailyIntake * 7
```



```{r Merge data sets}
# Select desired columns
# NDNS years 1 to 11
df_NDNS_seafoodReported_Yr1_11_short <- df_NDNS_seafoodReported_Yr1_11 %>%
  select(SurveyYear,CommonSpeciesName, avgUKdailyIntake, avgUKWeeklyIntake, SACN)

# Can add in species type here

#colnames(df_NDNS_seafoodReported_Yr1_11)
#colnames(df_NDNS_seafoodReported_Yr1_11_short)

# NDNS year 12
df_NDNS_seafoodReported_Yr12_short <- df_NDNS_seafood_UKReportedIntake_Yr12 %>%
  select(SurveyYear,CommonSpeciesName, avgUKdailyIntake, avgUKWeeklyIntake, SACN)

head(df_NDNS_seafoodReported_Yr12_short)

## Rename column
#df_NDNS_seafoodReported_Yr12_short <- df_NDNS_seafoodReported_Yr12_shorter %>% 
#  rename(
#    DayNo = RecallNo,
#    FoodName = FoodDescription)

# Combine years
df_NDNS_seafoodReported_Yr1_12 <- rbind(df_NDNS_seafoodReported_Yr1_11_short, df_NDNS_seafoodReported_Yr12_short)

```


Before merging with population data need to change survey Volumes. 

```{r Allocate numerical Volumes to survey years}
# Create new data frame with survey years and corresponding numerical Volumes
SurveyYear <- c("NDNS Year 1", "NDNS Year 2", "NDNS Year 3","NDNS Year 4","NDNS Year 5", "NDNS Year 6", "NDNS Year 7","NDNS Year 8","NDNS Year 9", "NDNS Year 10", "NDNS Year 11", "NDNS Year 12")
#NDNSYear <- c(200809, 200910, 201011, 201112, 201213, 201314, 201415, 201516, 201617, 201718, 201819) 
# 1.5 years. Choose starting or finishing year.
#Year <- c(2008, 2009, 2010, 2011, 2012, 2013, 2014, 2015, 2016, 2017, 2018) 

Year <- c(2009, 2010, 2011, 2012, 2013, 2014, 2015, 2016, 2017, 2018, 2019, 2020) 

# Create data frame with survey year and corresponding calendar year 
df_NDNSYearData <- as.data.frame(cbind(SurveyYear, Year))
df_NDNSYearData$Year <- as.numeric(as.character(df_NDNSYearData$Year))

# Create new data frame allocate numerical Volumes by merging data frames
#df_NDNS_seafood_consumption_200818 <- merge(df_NDNS_seafoodReported_Yr1_11, df_NDNSYearData, by = "SurveyYear")

# Omit year 1 i.e. 2008
#df_NDNS_seafood_consumption_200918 <- subset(df_NDNS_seafood_consumption_200818, Year > 2008)



# Create new data frame allocate numerical Volumes by merging data frames
df_NDNS_seafood_consumption_200920 <- merge(df_NDNS_seafoodReported_Yr1_12, df_NDNSYearData, by = "SurveyYear")

# Omit year 1 i.e. 2008
#df_NDNS_seafood_consumption_200920 <- subset(df_NDNS_seafood_consumption_200818, Year > 2008)

#unique(df_NDNS_seafood_consumption_200920$Year)

```



```{r Calcaulte seafood intake for whole of UK}
# Scale up NDNS consumption data with UK population data
# Join seafood consumption with population data by year
df_NDNS_seafoodintakeUK_200920 <- merge(df_NDNS_seafood_consumption_200920, df_UK_population_200920, by = "Year")

# Multiply by the number of people in the country
df_NDNS_seafoodintakeUK_200920$Volume <- (df_NDNS_seafoodintakeUK_200920$avgUKdailyIntake * 365) * df_NDNS_seafoodintakeUK_200920$UKPOP

```


```{r Add additional variables to data set}
# Create "cleaned" data frame, defining units and data source 
df_NDNS_seafoodintakeUK_200920$DataSupplier <- "NDNS"
df_NDNS_seafoodintakeUK_200920$DataSet <- "Food Diaries"
df_NDNS_seafoodintakeUK_200920$Commodity <- "Consumption"
df_NDNS_seafoodintakeUK_200920$Flag <- "ConV2"
df_NDNS_seafoodintakeUK_200920$TemporalResolution <- "CalendarYear"
df_NDNS_seafoodintakeUK_200920$Units <- "Grams"
df_NDNS_seafoodintakeUK_200920$FlagDescription <- "Unit converted from g/capita/week to g/country/year"

```


```{r Select desired variables for the data set}
# Identify columns in data frame
colnames(df_NDNS_seafoodintakeUK_200920)

# Select desired columns for cleaned data set
#df_NDNS_seafoodintakeUK_200920_cleaned <- df_NDNS_seafoodintakeUK_200920[c("DataSupplier", "DataSet", "Commodity", "SeafoodSpecies", "SpeciesType", "SACN", "Volume", "Units", "Year", "TemporalResolution", "Flag", "FlagDescription")]

df_NDNS_seafoodintakeUK_200920_cleaned <- df_NDNS_seafoodintakeUK_200920[c("DataSupplier", "DataSet", "Commodity", "CommonSpeciesName", "SACN", "Volume", "Units", "Year", "TemporalResolution", "Flag", "FlagDescription")]

# Rename "SeafoodSpecies" column 
df_NDNS_seafoodintakeUK_200920_clean <- rename(df_NDNS_seafoodintakeUK_200920_cleaned, Species = SeafoodSpecies)

df_NDNS_seafoodintakeUK_200920_clean <- rename(df_NDNS_seafoodintakeUK_200920_cleaned, Species = CommonSpeciesName)

```


```{r Check dataset}
# Check for duplicates
duplicates_NDNS <- df_NDNS_seafoodintakeUK_200920_clean %>% select(DataSupplier, DataSet, Commodity, Species, Year, Volume, Units) %>% group_by(DataSupplier, DataSet, Commodity, Species, Year, Volume, Units) %>% 
  filter(n()>1) %>% mutate(n = n())

# Check data transforms from long and thin to short and wide
shape_NDNS <- df_NDNS_seafoodintakeUK_200920_clean %>% 
  select(DataSupplier, DataSet, Commodity, Species, SACN, Year, Volume, Units) %>% 
  filter(Commodity %in% c("Consumption")) %>% 
  pivot_wider(names_from = Commodity, values_from = Volume)

```


```{r Save cleaned data}
# Save cleaned data as .csv
filepath <- paste(data_dir,"ProcessedData/", sep = "")
write.csv(df_NDNS_seafoodintakeUK_200920_clean, file = paste(filepath,"ConsumptionData_NDNS_PreliminaryCleaned.csv", sep = ""), row.names = FALSE)

```
