---
title: "Cleaning_NutrientComp-data_PHENOR"
author: "AL"
date: "`r Sys.Date()`"
output: html_document
---


```{r setup, include=FALSE}
# sets options for entire document in the first chunk
knitr::opts_chunk$set(echo = TRUE)
```

## Preparation

```{r Load libraries, include=FALSE}
# Load packages
library(tidyr) # for tidying data
library(dplyr) # base R alternative but neater
library(vroom) # for loading and transforming data
library(tidyverse) # data exploration 
library(reshape2) # for melting data frames i.e. short and wide to long and thin
library(data.table) # for fread()-function
library(mice) # md.pattern to show missing data
library(stringr) # used to replace matched patterns in a string
library(haven) #Read SAV files
library(readxl)
library(readr) # CSV file I/O, e.g. the read_csv function
library(sjlabelled)
library(compareDF)#Check missing values across datasets
library(chron)

```


```{r Load data, echo=TRUE}
# Loading file path to the data into a Volume 
source("Data_filepath.R")# Data_filepath.R is listed in .gitignore-file. So, you will need to create that file yourself and provide your respected filepath using "data_dir <- [enter your here]"
```




## Nutrient composition data (Public Health England)

Next we downloaded seafood nutrient composition data from Public Health England (PHE). This data includes the nutrient content (macro and micro nutrients) of commonly consumed fish in the UK. Composition analyses were conducted on raw and cooked fish samples.

Composition analyses were conducted in 2004. Temporal changes in nutrient composition is unlikely.

Note that .csv file is saved as a UT8 (otherwise an error appears. Check with Bernhard the differences in the file types)


```{r Specify the desired file path}
# Select file path with raw data
filepath <- paste(data_dir,"RawData/csv-files", sep = "") 
```


```{r Load PHE data (2004)}
# Read in .csv file 
df_PHE_nutreintComp <- vroom(file=paste(filepath,"PHE_seafoodNutrientContent.csv", sep="/"), na = c("N/A", "Tr")) 
# Both N/A and Tr different 

# Check for missing Volumes
#md.pattern(df_PHE_nutreintComp, rotate.names = TRUE)

# Identify column names
colnames(df_PHE_nutreintComp)

# Read in fatty acid data
df_PHE_FAComp <- vroom(file=paste(filepath,"PHE_seafoodFAContent.csv", sep="/"), na = c("N/A", "Tr")) 

# Create new column for omega-3 (sum EPA and DHA)
df_PHE_FAComp$Omega3_EPADHA_g100g <- df_PHE_FAComp$`cis n-3 C20:5_EPA` + df_PHE_FAComp$`cis n-3 C22:6_DHA`

# Join PHE nutrient composition data frame with fatty acid data frame 
df_PHE_nutreintCompFA <- df_PHE_nutreintComp %>% left_join(., df_PHE_FAComp, by = c('SampleDescription' = 'FoodGroup'))

```


```{r Load Norwegian data}
# Read in .csv file 
df_Norwegian_nutreintComp <- vroom(file=paste(filepath,"Norwegian_seafoodNutrientContent.csv", sep="/"), na = c("M", ""))

# Check for missing Volumes
#md.pattern(df_Norwegian_nutreintComp, rotate.names = TRUE)

# Identify column names
colnames(df_Norwegian_nutreintComp)

# Create new column for omega-3 (sum EPA and DHA)
df_Norwegian_nutreintComp$Omega3_EPADHA_g <- df_Norwegian_nutreintComp$`C20:5n-3_EPA_g` + df_Norwegian_nutreintComp$`C22:6n-3_DHA_g`

```


Select main nutrients. Both PHE and Norwegian datasets in g/100g.

```{r Select desired variables from both datasets}
# Select desired variables
df_PHE_nutreintComp_short <- df_PHE_nutreintCompFA[c("SampleDescription", "Protein-g100g", "Omega3_EPADHA_g100g", "VitaminA-µg100g", "VitaminD-µg100g", "VitaminB12-µg100g", "Zinc-mg100g", "Iron-mg100g", "Calcium-mg100g", "Selenium-µg100g", "Iodine-µg100g" )]

# Select desired variables
df_Norwegian_nutreintComp_short <- df_Norwegian_nutreintComp[c("FoodItem", "Protein_g", "Omega3_EPADHA_g", "VitaminA_RAE", "VitaminD_µg", "VitaminB12_µg", "Zinc_mg", "Iron_mg", "Calcium_mg", "Selenium_µg", "Iodine_µg")]

```


```{r Rename variables and join datasets}
# Rename variables
names(df_PHE_nutreintComp_short) <- c("FoodItem", "Protein_g100g", "Omega3_DHAEPA_g100g", "VitaminA_µg100g", "VitaminD_µg100g", "VitaminB12_µg100g", "Zinc_mg100g", "Iron_mg100g", "Calcium_mg100g", "Selenium_µg100g", "Iodine_µg100g")

names(df_Norwegian_nutreintComp_short) <- c("FoodItem", "Protein_g100g", "Omega3_DHAEPA_g100g", "VitaminA_µg100g", "VitaminD_µg100g", "VitaminB12_µg100g", "Zinc_mg100g", "Iron_mg100g", "Calcium_mg100g", "Selenium_µg100g", "Iodine_µg100g")

# Combine data frames
df_all_nutrientComp <- rbind(df_PHE_nutreintComp_short,
                              df_Norwegian_nutreintComp_short) 

str(df_all_nutrientComp)
```



```{r Transform dataset}
#Change structure of the data set 
df_all_nutrient_long <- melt(df_all_nutrientComp, id.vars = c("FoodItem"), variable.name ="Nutrient", value.name = "Volume")

```


```{r Specify the desired file path}
filepath <- paste(data_dir,"Methods/SpeciesTypeClassification/SeafoodNutrientComposition", sep = "") 
# note we are able to change this to another file e.g. saving the output
```


```{r Load translation document}
# Load database 
df_seafoodCompKey <- vroom(file=paste(filepath,"SeafoodComposition_translationKey.csv", sep="/"), na = c("NA"))

str(df_seafoodCompKey)
```



```{r Join data with translation document}
# Join data frame with translation document- ensures consistent spelling across data sets

# Duplicate species name column (allows us to check species have been renames correctly)
df_all_nutrient_long$SpeciesRevised <- df_all_nutrient_long$FoodItem  

str(df_all_nutrient_long)

# Join data frame with MCS data frame by species name
df_all_nutrient_long <- df_all_nutrient_long %>% inner_join(., df_seafoodCompKey, by = c('SpeciesRevised' = 'MCS (DO NOT TOUCH)'))
# MSC (DO NOT TOUCH) column is the unique spelling of the species in all data

```


```{r Add additional variables to data set}
# Add additional variables
df_all_nutrient_long$DataSet <- "NutrientAnalysis"
df_all_nutrient_long$Commodity <- "NutrientComposition"
df_all_nutrient_long$Flag <- "NA"
df_all_nutrient_long$FlagDescription <- "NA"
df_all_nutrient_long$TemporalResolution <- "NA"
df_all_nutrient_long$Units <- "weight/100g"
df_all_nutrient_long$Year <- "NA"
```


```{r Select desired variables for the data set}
# Identify columns in data frame
colnames(df_all_nutrient_long)

# Select desired columns for cleaned data set
df_all_nutrient_cleaned <- df_all_nutrient_long[c("DataSupplier", "DataSet", "Commodity", "SeafoodSpecies", "SpeciesType", "SACN", "Nutrient", "Volume", "Units", "Year", "TemporalResolution", "Flag", "FlagDescription")]

# Rename variable
df_all_nutrient_clean <- rename(df_all_nutrient_cleaned, 
                                                    Species = SeafoodSpecies)

str(df_all_nutrient_clean)
#df_all_nutrient_clean <- df_all_nutrient_cleaned %>% rename("Species" == "SeafoodSpecies")
#df_all_nutrient_clean <- df_all_nutrient_cleaned %>% rename("SeafoodSpecies" == "Species")
#df_all_nutrient_clean <- df_all_nutrient_cleaned %>% rename(SeafoodSpecies = Species)

```


```{r Save cleaned data}
# Change file path to "ProcessedData"
filepath <- paste(data_dir,"ProcessedData/", sep="")
write.csv(df_all_nutrient_cleaned, file =
            paste(filepath,"NutrientComp_PHENor_PreliminaryCleaned.csv", sep=""), row.names = FALSE)

```
