---
title: "RI-B5-04: UK seafood supply chains. Database assembly."
author: "B. Scheliga and A.Lofstedt"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

This is a document describes the methods and data used to create the UK Seafood supply chain database, linking  
both capture and aquaculture), trade, consumption and purchasing behavior between 2009 and 2018.

This project was part of the RESAS - Project from the Rowett Institute at the University of Aberdeen.

Two data frames are created- one with seafood supplies expressed in grams and another expressed in g/ capita/ week 


## Preparation

```{r Load libraries}
library(tidyr) # for tidying data
library(dplyr) # base R alternative but neater
library(vroom) # for loading and transforming data
library(tidyverse) # data exploration 
library(reshape2) # for melting data frames i.e. short and wide to long and thin
library(data.table) # for fread()-function
library(mice) # md.pattern to show missing data
library(stringr) # used to replace matched patterns in a string

```


```{r Load cleaned data}
# loading filepath to the project folder on the University storage
source("Data_filepath.R")

# Only selecting the cleaned dataset, which are annotate with the "_cleaned_sfdb". There might be other processed files in the folder, which we don't want
filepath <- paste(data_dir,"ProcessedData/", sep="")# we are storing the file path here, because we'll need it later
vec_filenames_sfdb <- list.files(filepath, pattern = "Cleaned.csv", full.names = TRUE)


print("The following files were loaded:")
i = 1
# Loading the cleaned dataset
for(f in vec_filenames_sfdb){
    temp <- fread(f) # storing the data in a temporary 
    assign(paste("df_",substr(sub(paste(".*",filepath,sep=""),"",f),1,6),sep=""),temp) # assigning the df a name based on input dataset
    print(paste(i,". ",sub(paste(".*",filepath,sep=""),"",f),sep="")) # print out which dataset files had been used
    i = i+1
    rm(temp)# removing temp object
}
rm(i)

```


## A) Create seafood database (g)

```{r rbind all cleaned datasets}
#df_Seafood_DB <- rbind(df_Trade_, df_AllLan)# Note df_Trade_ is the edited one

df_Seafood_DB <- rbind(df_TradeD, df_UKVess) 
df_Seafood_DB <- rbind(df_Seafood_DB, df_Aquacu)
df_Seafood_DB <- rbind(df_Seafood_DB,df_Consum)
df_Seafood_DB <- rbind(df_Seafood_DB,df_EatenO)
df_Seafood_DB <- rbind(df_Seafood_DB,df_Househ)

# Check the structure
str(df_Seafood_DB)
```


```{r Save the seafood database}
source("Data_filepath.R")
# Save datframe to "Outputs" data file
filepath <- paste(data_dir,"Outputs/Databases", sep="")
write.csv(df_Seafood_DB, paste(filepath,"Seafood-database_preliminary.csv",sep="/"), row.names = FALSE)
```



# B) Seafood database (g/capita/week)

```{r Specify the desired file path}
filepath <- paste(data_dir,"Outputs/Databases", sep = "") 
# note we are able to change this to another file e.g. saving the output
```

```{r Load the seafood database}
# Load database 
df_Seafood_DB <- vroom(file=paste(filepath,"Seafood-database_preliminary.csv", sep="/"), na = c("NA"))
```


```{r Specify the desired file path}
filepath <- paste(data_dir,"ProcessedData", sep = "") 
```

```{r Load population data}
# Load database 
df_UK_population <- vroom(file=paste(filepath,"UKPopulation_ONS_PreliminaryCleaned.csv", sep="/"), na = c("NA"))
```



```{r Change struucture of the dataframes}
# Simplify population data
colnames(df_UK_population)
df_UK_population_simp <- df_UK_population[c("Year", "UKPOP")]

# Simplify data set and select desired columns 
df_UK_SeafoodDB_simp <- df_Seafood_DB[c("Species", "SpeciesType", "SACN", "Year", "Commodity", 
                                                    "Volume" )]

# Aggregate value by commodity, year and species (production data composed of aquaculture and wild capture)
df_UK_SeafoodDB_simp_agg <- aggregate(Volume ~ Commodity + Year + Species + SpeciesType + SACN, data = df_UK_SeafoodDB_simp, sum)

# Need to unstack database to calculate seafood supplied. Change structure of the dataframe- convert to wide and short
df_UK_SeafoodDB_short <- df_UK_SeafoodDB_simp_agg %>%
  pivot_wider(names_from = Commodity, values_from = Volume)
str(df_UK_SeafoodDB_short)

# Convert NA to 0
df_UK_SeafoodDB_short[is.na(df_UK_SeafoodDB_short)] <- 0

# Calculate supplies
df_UK_SeafoodDB_short$Supplies <- (df_UK_SeafoodDB_short$Production + df_UK_SeafoodDB_short$Imports) - df_UK_SeafoodDB_short$Exports


# Merge data frame with population data
df_seafoodDB_pop_short <- merge(df_UK_SeafoodDB_short, df_UK_population_simp, "Year")
str(df_seafoodDB_pop_short)

df_seafoodDB_pop_short$Production_pcw <- (df_seafoodDB_pop_short$Production/
                                            df_seafoodDB_pop_short$UKPOP)/52
df_seafoodDB_pop_short$Imports_pcw <- (df_seafoodDB_pop_short$Imports/
                                            df_seafoodDB_pop_short$UKPOP)/52
df_seafoodDB_pop_short$Exports_pcw <- (df_seafoodDB_pop_short$Exports/
                                            df_seafoodDB_pop_short$UKPOP)/52
df_seafoodDB_pop_short$Supplies_pcw <- (df_seafoodDB_pop_short$Supplies/
                                            df_seafoodDB_pop_short$UKPOP)/52
df_seafoodDB_pop_short$Consumption_pcw <- (df_seafoodDB_pop_short$Consumption/
                                            df_seafoodDB_pop_short$UKPOP)/52
```


```{r Save the seafood database}
source("Data_filepath.R")
# Save datframe to "Outputs" data file
filepath <- paste(data_dir,"Outputs/Databases", sep="")
write.csv(df_seafoodDB_pop_short, paste(filepath,"Seafood-database-PCW_preliminary.csv",sep="/"), row.names = FALSE)
```





## C) Create database with nutrients

```{r Specify the desired file path}
filepath <- paste(data_dir,"ProcessedData", sep = "") 
```

```{r Load nutrient data}
# Load database 
df_PHE_nutrient <- vroom(file=paste(filepath,"NutrientComp_PHENor_PreliminaryCleaned.csv", sep="/"), na = c("NA"))
```

```{r Simplify nutrient dataframe and unstack}
# Subset desired columns
df_PHE_nutrient_simp <- df_PHE_nutrient[,c("Species", "Nutrient", "Volume")]

# Unstack nutrients data 
# Make data frame wide and short
df_nutrient_short <- df_PHE_nutrient_simp %>%
  pivot_wider(names_from = Nutrient, values_from = Volume)

str(df_nutrient_short)
```

```{r Change struucture of the dataframes}
# Merge nutrient data with seafood database
df_seafoodDB_PCW_nutreintComposition <- merge(df_seafoodDB_pop_short, df_nutrient_short, "Species") 
# Only use this dataframe when looking at nutrients as all fish are not included
str(df_seafoodDB_PCW_nutreintComposition)

```


Create seafood database where commodities are unstacked

```{r Save the seafood database}
source("Data_filepath.R")
# Save datframe to "Outputs" data file
filepath <- paste(data_dir,"Outputs/Databases", sep="")
write.csv(df_seafoodDB_PCW_nutreintComposition, paste(filepath,"Seafood-database-PCW_composition_preliminary.csv",sep="/"), row.names = FALSE)
```




