---
title: "Part 1. Data cleaning. RI-B5-04: UK seafood supply chains"
author: "A.Lofstedt, B.Scheliga"
date: '17/01/2023'
output: html_document
---

```{r setup, include=FALSE}
# sets options for entire document in the first chunk
knitr::opts_chunk$set(echo = TRUE)
```

## Part 1: Cleaning data used to map UK seafood supply chains 

This R markdown document outlines how the data compiled in the UK seafood database were cleaned. Data on seafood production (both capture and aquaculture), trade, consumption and purchasing behavior between 2009 and 2018 were sourced. 

Justification for the data sources used can be found in the supporting excel document. Prior to loading the data into R, data were preliminary cleaned to allow compatibility with R. Details describing this process can be foundd in the supporting word document. 

## Preparation

```{r Load libraries, include=FALSE}
# Load packages
library(tidyr) # for tidying data
library(dplyr) # base R alternative but neater
library(vroom) # for loading and transforming data
library(tidyverse) # data exploration 
library(reshape2) # for melting data frames i.e. short and wide to long and thin
library(data.table) # for fread()-function
library(mice) # md.pattern to show missing data
library(stringr) # used to replace matched patterns in a string

```


```{r Load data, echo=TRUE}
# Loading file path to the data into a value 
source("Data_filepath.R")# Data_filepath.R is listed in .gitignore-file. So, you will need to create that file yourself and provide your respected filepath using "data_dir <- [enter your here]"
```


```{r Specify the desired file path}
filepath <- paste(data_dir,"Methods/SpeciesTypeClassification", sep="") 
# note we are able to change this to another file e.g. saving the output
```


```{r Load spreadsheet with unique seafood species translations}
# Load master spreadsheet with species names and revised species names
MCS_translations <- vroom(file=paste(filepath,"AllMainCommercialSpecies_translations.csv", sep="/"))

# Check structure of the translation document
str(MCS_translations)
```


```{r Specify the desired file path}
# Select file path with raw data
filepath <- paste(data_dir,"RawData/csv-files", sep = "") 
```


## Production

UK seafood production data includes both data from capture fisheries and aquaculture production. Multiple data sets are loaded to accommodate for the time period.

### A) Capture fisheries (Office of National Statistics)

```{r Load data (2017 to 2020)}
# i) Landings into the UK by UK vessels: 2017 to 2021
# Read in .csv file
df_UK_landings_17_21_data <- vroom(file=paste(filepath,"NS_UK_landings-201721.csv", sep="/"))

# Check for missing values
md.pattern(df_UK_landings_17_21_data, rotate.names = TRUE)

# Select desired variables and years (2017 to 2020)
df_UK_landings_17_20_data <- df_UK_landings_17_21_data[c("SpeciesType","Species", "2017", "2018","2019", "2020")]
```


```{r Transform data set (2017 to 2020)}
# Change structure of the data set 
df_UK_landings_201720_long <- melt(df_UK_landings_17_20_data, id.vars = c("SpeciesType", "Species"), variable.name = "Year", value.name = "Value")
```


```{r Load data (2012 to 2016)}
# ii) Landings into the UK by UK vessels: 2012 to 2016 
# Read in .csv file
df_UK_landings_12_16_data <- vroom(file=paste(filepath,"NS_UK_landings-201216.csv", sep="/"))

# Check for missing values
md.pattern(df_UK_landings_12_16_data, rotate.names = TRUE)

```


```{r Transform data set (2012 to 2016)}
# Change structure of the data set 
df_UK_landings_201216_long <- melt(df_UK_landings_12_16_data, id.vars = c("SpeciesType", "Species"), variable.name = "Year", value.name = "Value")
```


```{r Load data (2009 to 2011)}
# iii) Landings into the UK by UK vessels: 2008 to 2012 but select years 2009 to 2011. Always best to use the most recent data set.
# Read in .csv file
df_UK_landings_08_12_data <- vroom(file=paste(filepath,"NS_UK_landings-200812.csv", sep="/"))

# Check for missing values
md.pattern(df_UK_landings_08_12_data, rotate.names = TRUE)

# Check column names 
colnames(df_UK_landings_08_12_data)

# Select desired variables and years (2009 to 2011)
df_UK_landings_09_11_data <- df_UK_landings_08_12_data[c("SpeciesType","Species", "2009", "2010","2011")]

```


```{r Transform data set (2009 to 2011)}
# Change structure of the data set 
df_UK_landings_200911_long <- melt(df_UK_landings_09_11_data, id.vars = c("SpeciesType", "Species"), variable.name="Year", value.name="Value")

```


```{r Combine data sets: i) 2009 to 2011; ii) 2012 to 2016; and iii) 2017 to 2020)}
# Combine capture production data i) 2009 to 2011; ii) 2012 to 2016; and iii) 2017 to 2020)
df_UK_landings_09_20 <- rbind(df_UK_landings_200911_long,
                              df_UK_landings_201216_long, 
                              df_UK_landings_201720_long)
```


Capture fisheries data are now all in one data frame. Data frame is composed of four columns: SpeciesType, Species, Year and Value. 

We need to ensure species are spelled the same across all data sets, otherwise data sets cannot be merged, for instance "Blue whiting" v.s. "Blue Whiting or "Crab" v.s. "Crabs". We created an excel document listing the unique spellings of all species in all data sets used. In the cleaning process, the data sheets are joined to the "translation" document by the unique species name. In the final cleaned version, the correct species name column is selected. The translation document also provides various species type classifications as per published peer-reviewed papers.


```{r Join data frame with translation document- ensures consistent spelling across data sets}
# Duplicate species name column (allows us to check species have been renames correctly)
df_UK_landings_09_20$SpeciesNEW <- df_UK_landings_09_20$Species

# Join data frame with MCS data frame by species name
df_UK_landings_09_20 <- df_UK_landings_09_20 %>% inner_join(., MCS_translations, by = c('SpeciesNEW' = 'MCS (DO NOT TOUCH)'))
# MSC (DO NOT TOUCH) column is the unique spelling of the species in all data sets.

# Check column names of the updated data frame
colnames(df_UK_landings_09_20)

```



Landings provided as whole weight. Need to apply conversion factors estimate edible portion of the seafood. Converstion factors taken from Lofstedt et al 2021. 

```{r Apply conversion factors to calculate edible portion of seafood}
# Classify species as finfish or shellfish. New column for conversion name
unique(df_UK_landings_09_20$FAO_ST)

# Specify shellfish
df_UK_landings_09_20$CN[df_UK_landings_09_20$FAO_ST %in% c("Crustaceans", "Molluscs", "Cephalopods", "Other molluscs and aquatic invertebrates", "Other shellfish")] <- "Shellfish"

# Specify finfish
df_UK_landings_09_20$CN[df_UK_landings_09_20$FAO_ST %in% c("Freshwater fish", "Demersal fish", "Other marine fish", "Pelagic fish")] <- "Finfish"

# Create a new column stating the conversion factor. Conversion factors used to convert whole fish weight to edible weight
# Average conversion factors for finfish: 0.49; shellfish: 0.28 (Lofstedt et al., 2021)

# State average CF finfish: 0.49; shellfish: 0.28 
df_UK_landings_09_20$CF[df_UK_landings_09_20$CN == "Shellfish"] <- 0.28
df_UK_landings_09_20$CF[df_UK_landings_09_20$CN == "Finfish"] <- 0.49

# Create new column and apply conversion factors values
df_UK_landings_09_20$CFValue <- df_UK_landings_09_20$Value * df_UK_landings_09_20$CF

```


```{r Convert from 1000 tonnes to grams}
# Convert raw value from 000's tonnes to grams 
df_UK_landings_09_20$Value <- (df_UK_landings_09_20$CFValue * 1000) * 1000000

```


```{r Add additional variables to data set}
# Create "cleaned" data frame, defining units and data source 
df_UK_landings_09_20$Units <- "Grams"
df_UK_landings_09_20$DataSupplier <- "NationalStatistics"
df_UK_landings_09_20$DataSet <- "UK_SeaFisheriesStatistics"
df_UK_landings_09_20$Commodity <- "CaptureProduction"
df_UK_landings_09_20$Flag <- "ConV1, ConV2"
df_UK_landings_09_20$TemporalResolution <- "Annual"

```


```{r Select desired variables for the cleaned data set}
# Identify column names
colnames(df_UK_landings_09_20)

# Select desired columns for cleaned data set-- 16  variables
df_cleaned_UK_landings_200920 <- df_UK_landings_09_20[c("RevisedMCS", "Year", "Value", "Units", "EUMOFA_ST", "FAO_ST", "ISSCAAP_ST", "Golden_2021_ST","Bianchi_2022_ST", "WWF_2022_ST", "SACN", "DataSupplier", "DataSet", "Commodity", "Flag", "TemporalResolution")]

```


```{r Save cleaned data}
# Change file path to "ProcessedData". Can also use setwd()
filepath <- paste(data_dir,"ProcessedData", sep = "")
write.csv(df_cleaned_UK_landings_200920, file = paste(filepath,"CaptureProductionData_NS_clean.csv", sep = ""))

```



### B) Aquaculture production (EUROSTAT)

```{r Load data (2009 to 2018)}
# Read in .csv file 
df_EUROSTAT_UK_aquaculture_data <- vroom(file=paste(filepath,"EUROSTAT_UK_aquaculture-200918-edit.csv", sep="/"),na = c(":"))

# Check for missing values
md.pattern(df_EUROSTAT_UK_aquaculture_data, rotate.names = TRUE)

# Check column names
colnames(df_EUROSTAT_UK_aquaculture_data)

```


```{r Transform data set (2009 to 2018)}
# Change structure of the data set. id.vars states the number of columns not to be stacked
df_EUROSTAT_UKFarmed <- melt(df_EUROSTAT_UK_aquaculture_data, id.vars = c("Species"), variable.name = "Year", value.name = "Value")

```


```{r Rename selected individual species and provide species types}
# Duplicate species name column (duplicated column entries to be duplicated)
df_EUROSTAT_UKFarmed$SpeciesNEW <- df_EUROSTAT_UKFarmed$Species

# Merge data frame with MCS data frame by species name
df_EUROSTAT_UKFarmed <- df_EUROSTAT_UKFarmed %>% inner_join(., MCS_translations, by = c('SpeciesNEW' = 'MCS (DO NOT TOUCH)'))

```


```{r Apply conversion factor to calculate edible portion of seafood}
# Classify species as finfish or shellfish using FAO species type columm. New column for conversion name

# Check unique FAO species type commodities 
unique(df_EUROSTAT_UKFarmed$FAO_ST)


# Specify shellfish
df_EUROSTAT_UKFarmed$CN[df_EUROSTAT_UKFarmed$FAO_ST %in% c("Crustaceans", "Molluscs", "Other molluscs and aquatic invertebrates")] <- "Shellfish"

# Specify finfish
df_EUROSTAT_UKFarmed$CN[df_EUROSTAT_UKFarmed$FAO_ST %in% c("Freshwater fish", "Demersal fish", "Other marine fish")] <- "Finfish"

# Create a new column stating the conversion factor. Conversion factors used to convert whole fish weight to edible weight
# Average conversion factors for finfish: 0.49; shellfish: 0.28 (Lofstedt et al., 2021)
df_EUROSTAT_UKFarmed$CF[df_EUROSTAT_UKFarmed$CN == "Shellfish"] <- 0.28
df_EUROSTAT_UKFarmed$CF[df_EUROSTAT_UKFarmed$CN == "Finfish"] <- 0.49

# Create new column and apply conversion factors values
df_EUROSTAT_UKFarmed$CFValue <- df_EUROSTAT_UKFarmed$Value * df_EUROSTAT_UKFarmed$CF

```


```{r Convert from tonnes to grams}
# Convert raw value from tonnes to grams 
df_EUROSTAT_UKFarmed$Value <- df_EUROSTAT_UKFarmed$CFValue * 1000000

```


```{r Add additional variables to data set}
# Create "cleaned" data frame, defining units and data source 
df_EUROSTAT_UKFarmed$DataSupplier <- "Eurostat"
df_EUROSTAT_UKFarmed$DataSet <- "Aquaculture"
df_EUROSTAT_UKFarmed$Commodity <- "Production"
df_EUROSTAT_UKFarmed$Flag <- "Conv2"
df_EUROSTAT_UKFarmed$Units <- "Grams"
df_EUROSTAT_UKFarmed$TemporalResolution <- "Annual"

```


```{r Select desired variables for the data set}
# Identify column names
colnames(df_EUROSTAT_UKFarmed)

# Select desired columns for cleaned data set
df_cleaned_UK_aquaculture_200918 <- df_EUROSTAT_UKFarmed[c("RevisedMCS", "Year", "Value", "Units", "EUMOFA_ST", "FAO_ST", "ISSCAAP_ST", "Golden_2021_ST","Bianchi_2022_ST", "WWF_2022_ST", "SACN", "DataSupplier", "DataSet", "Commodity", "Flag")]

```


```{r Save cleaned data}
# Change file path to "ProcessedData"
filepath <- paste(data_dir,"ProcessedData", sep = "")
write.csv(df_cleaned_UK_aquaculture_200918, file = 
            paste(filepath,"AquacultureData_Eurostat_clean.csv", sep = ""))

```


## Delete later

```{r DELETE LATER. Check to see if species columns successfully merge}

# Compare column names 
colnames(df_cleaned_UKFarmed)
colnames(df_cleaned_UK_landings_200920)

# Merge data frames by species name and year
check_names <- merge(df_cleaned_UK_landings_200920, df_cleaned_UKFarmed, 
                     by = c("SpeciesNEW", "Year"))

# Check the species names that overlap between the two data sets
unique(check_names$SpeciesNEW)
unique(df_cleaned_UK_landings_200920$SpeciesNEW)
unique(df_cleaned_UKFarmed$SpeciesNEW)

```


## Trade (HMRC)

Trade data was sourced from the HMRC and includes both imports and exports. An updated script for cleaning the trade data can be found in a separate file. 

```{r Load data from processed data file}
# Some edits are needed to ensure consistency with processed data sets

# Read in UK trade data 
df_trade_HMRC_data <- vroom(file=paste(filepath,"TradeData_HMRC_Preliminary-Cleaned.csv", sep = "/"))
str(df_trade_HMRC_data)

```

```{r Rename selected individuial species}
# Duplicate species name column (duplicated column entries to be duplicated)
df_trade_HMRC_data$SpeciesNEW <- df_trade_HMRC_data$Species

# Merge data frame with MCS data frame by species name
df_trade_HMRC_data <- df_trade_HMRC_data %>% inner_join(., MCS_translations, by = c('SpeciesNEW' = 'MCS (DO NOT TOUCH)'))

```


```{r OLD Calculate grams available}
# Create new column converting values from '000 tonnes to grams
df_trade_HMRC_data$ValueGrams <- (df_trade_HMRC_data$Value*1000)*1000000

```


```{r Select desired values}
# Identify columns in data frame
colnames(df_trade_HMRC_data)

# Select desired columns for cleaned data set
df_cleaned_UK_HMRC_200919 <- df_trade_HMRC_data[c("RevisedMCS", "Year", "Value", "Units", "EUMOFA_ST", "FAO_ST", "ISSCAAP_ST", "Golden_2021_ST","Bianchi_2022_ST", "WWF_2022_ST", "SACN", "DataSupplier", "DataSet", "Commodity", "Flag")]

```


```{r Save cleaned data}
# Save cleaned data as .csv
filepath <- paste(data_dir,"ProcessedData", sep = "")
write.csv(df_cleaned_UK_HMRC_200919, file = paste(filepath,"Trade_HMRC_clean.csv", sep = ""))

```


## Delete later

```{r DELETE LATER. Check to see if species columns successfully merge}
# Compare column names 
colnames(df_cleaned_UKFarmed)
colnames(df_cleaned_UK_landings_200920)
colnames(df_clean_trade_HMRC)

# Merge data frames by species name and year
check_names_again <- merge(df_cleaned_UK_landings_200920, df_clean_trade_HMRC, 
                     by = c("SpeciesNEW", "Year"))

# to check: monk, pollock, 

# Check the species names that overlap between the two data sets
unique(check_names_again$SpeciesNEW)
unique(df_cleaned_UK_landings_200920$SpeciesNEW)
unique(df_clean_trade_HMRC$SpeciesNEW)
```



## UK seafood consumption (to food level) (NDNS)

UK fish consumption data was sourced from the National Diet and Nutrition Survey (NDNS). Food diaries show the amount of seafood and often species consumed in the UK. It has greater granularity compared to other data sources (consumption identified to species level).

UK seafood consumption data is collected on a biannual basis i.e. participants provide food diary entries every two years. Although food diary entry dates are provided, yearly consumption per calendar year cannot be calculated due to the nature of the data collection. Therefore, over a 10-year period, there will be five seafood consumption estimates. NDNS data is freely available from 2008 to 2019 (i.e. reporting years 1 to 11). Data, to food level consumption can be downloaded as txt files from the UK data service.

Convert all tab files to .csv before loading into R


```{r Load data}
# Load NDNS yrs 1 - 4
df_NDNS_foodlevel_yr1_4_data <- vroom(file= paste(filepath,"NDNS_UK_FoodLevel-Y1-4.csv", sep = "/"))

# Load NDNS yrs 5 - 6 
df_NDNS_foodlevel_yr5_6_data <-
  vroom(file = paste(filepath,"NDNS_UK_FoodLevel-Y5-6.csv", sep = "/"))

# Load NDNS yrs 7 - 8 
df_NDNS_foodlevel_yr7_8_data <-
  vroom(file = paste(filepath,"NDNS_UK_FoodLevel-Y7-8.csv", sep = "/"))

# Load NDNS yrs 9-11
df_NDNS_foodlevel_yr9_11_data <- 
  vroom(file = paste(filepath,"NDNS_UK_FoodLevel_Y9-11.csv", sep = "/"))

```


```{r Join food level NDNS data together in one data frame}
# Check column names
colnames(df_NDNS_foodlevel_yr1_4_data)
colnames(df_NDNS_foodlevel_yr5_6_data)
colnames(df_NDNS_foodlevel_yr7_8_data)
colnames(df_NDNS_foodlevel_yr9_11_data)

# Omit "DiaryDate" column from the yr9-11 data set
df_NDNS_foodlevel_yr9_11_data_clean <- subset(df_NDNS_foodlevel_yr9_11_data, select = -c(DiaryDate))

# Rename "AgeR"
colnames(df_NDNS_foodlevel_yr9_11_data_clean)[colnames(df_NDNS_foodlevel_yr9_11_data_clean) == 'AgeR'] <- 'Age'

# Check column was omitted 
colnames(df_NDNS_foodlevel_yr9_11_data_clean)

# Join data frames
NDNS_UK_Yr1_11_data <- rbind(df_NDNS_foodlevel_yr1_4_data,
                             df_NDNS_foodlevel_yr5_6_data,
                             df_NDNS_foodlevel_yr7_8_data,
                             df_NDNS_foodlevel_yr9_11_data_clean)

# Check for missing values
md.pattern(NDNS_UK_Yr1_11_data)

```


Downloaded NDNS data for all food groups. Need to select just those rows related to seafood consumption. Entries numbered 33, 34 and 35 "Main food group code" column related to seafood consumption.

```{r Subset seafood commodities from consumption data}
# Identify unique main food groups
unique(NDNS_UK_Yr1_11_data$MainFoodGroupCode)

# Subset seafood commodities. Seafood commodities identified as values: 33, 34 and 35 in "main food group code" column
df_NDNS_seafood_consumption_Yr1_11 <- subset(NDNS_UK_Yr1_11_data, MainFoodGroupCode == 33 | NDNS_UK_Yr1_11_data$MainFoodGroupCode == 34| NDNS_UK_Yr1_11_data$MainFoodGroupCode == 35)

# Check structure of seafood data frame
str(df_NDNS_seafood_consumption_Yr1_11)

# Create list of unique seafood items consumed
#seafoodItems_Yr1_11 <- unique(df_NDNS_seafood_consumption_Yr1_11$FoodName)

# Export as .csv
#write.csv(seafoodItems_Yr1_11, file = paste(filepath,"UniqueSeafoodItems.csv", sep = ""))
```


```{r Calculate total grams of seafood consummed}
# Calculate total seafood consumption. Sum columns"WhiteFishg","OilyFishg", "CannedTunag" and "Shellfishg"
df_NDNS_seafood_consumption_Yr1_11$Value <- df_NDNS_seafood_consumption_Yr1_11$WhiteFishg + df_NDNS_seafood_consumption_Yr1_11$OilyFishg + df_NDNS_seafood_consumption_Yr1_11$CannedTunag + df_NDNS_seafood_consumption_Yr1_11$Shellfishg

```


```{r Correct for differences in species spellings}
# Duplicate species name column (allows us to check species have been renames correctly)
df_NDNS_seafood_consumption_Yr1_11$SpeciesNEW <- df_NDNS_seafood_consumption_Yr1_11$FoodName

# Join data frame with MCS data frame by species name
df_NDNS_seafood_consumption_Yr1_11 <- df_NDNS_seafood_consumption_Yr1_11 %>% inner_join(., MCS_translations, by = c('SpeciesNEW' = 'MCS (DO NOT TOUCH)'))
# MSC (DO NOT TOUCH) column is the unique spelling of the species in all data sets.
```


```{r Check survey year data and data frame column names}
# Identify the survey years
unique(df_NDNS_seafood_consumption_Yr1_11$SurveyYear)

# Rename survey years to numerical values 
200809, 201011, 201213, 201415, 201617, 201819  

# Ensure column names are consistent between data sets
# Change "SurveyYear" to "Year"

# Change "TotalGrams" to "Value"



```




```{r Select desired variables}
# Identify columns in data frame
colnames(df_NDNS_seafood_consumption_Yr1_11)

# Select desired columns for cleaned data set
df_NDNS_seafood_consumption_Yr1_11_cleaned <- df_NDNS_seafood_consumption_Yr1_11[c("RevisedMCS", "Year", "Value",  "EUMOFA_ST", "FAO_ST", "ISSCAAP_ST", "Golden_2021_ST","Bianchi_2022_ST", "WWF_2022_ST", "SACN", "DataSupplier", "DataSet", "Commodity", "Flag")]

# Check what variable we need- all fish "WhiteFishg","OilyFishg", "CannedTunag","Shellfishg"?

```


```{r Add additional variables to data set}
# Create "cleaned" data frame, defining units and data source 
df_NDNS_seafood_consumption_Yr1_11$DataSupplier <- "NDNS"
df_NDNS_seafood_consumption_Yr1_11$DataSet <- "FoodDiaries"
df_NDNS_seafood_consumption_Yr1_11$Commodity <- "Consumption"
df_NDNS_seafood_consumption_Yr1_11$Flag <- "ConV1, ConV2"
df_NDNS_seafood_consumption_Yr1_11$TemporalResolution <- "Biennial"
df_NDNS_seafood_consumption_Yr1_11$Units <- "Grams"

```



```{r Save cleaned data}
# Save cleaned data as .csv
filepath <- paste(data_dir,"ProcessedData", sep = "")
write.csv(df_NDNS_seafood_consumption_Yr1_11_cleaned, file = paste(filepath,"ConsumptionData_NDNS_clean.csv", sep = ""))

```



## UK seafood purchases (DEFRA family food)

```{r Load data 2016 to 2020} 
# *Error in raw data in 2015, so data from 2016 onwards used

# read DEFRA purchasing (household and eating out) data (.csv)
df_DEFRA_purchase_201620_data <- vroom(file=paste(filepath,"DEFRA_UK_purchases-201620.csv", sep="/"))

md.pattern(df_DEFRA_purchase_201620_data)

# State full calendar year date
df_DEFRA_purchase_201620_data$Year[df_DEFRA_purchase_201620_data$Year==15] <- 2015
df_DEFRA_purchase_201620_data$Year[df_DEFRA_purchase_201620_data$Year==16] <- 2016
df_DEFRA_purchase_201620_data$Year[df_DEFRA_purchase_201620_data$Year==17] <- 2017
df_DEFRA_purchase_201620_data$Year[df_DEFRA_purchase_201620_data$Year==18] <- 2018
df_DEFRA_purchase_201620_data$Year[df_DEFRA_purchase_201620_data$Year==19] <- 2019
df_DEFRA_purchase_201620_data$Year[df_DEFRA_purchase_201620_data$Year==20] <- 2020
unique(df_DEFRA_purchase_201620_data$Year)
```


```{r Separate purchasing and eating out data} 
# Separate purchasing data for household and eating out purposes based on code. Create a pipe aggregating fish products consumed within the household and eaten out
df_DEFRA_eaten_out_201620 <- df_DEFRA_purchase_201620_data[df_DEFRA_purchase_201620_data$codecat %in% c( "eo120101", "eo120102", "eo120201", "eo120202", "eo1203", "eo1204", "eo1205", "eo120601", "eo120602", "eo120603"), ]

df_DEFRA_household_201620 <- df_DEFRA_purchase_201620_data[df_DEFRA_purchase_201620_data$codecat %in% c( "10201", "10202", "10601", "10602", "10701", "10702", "10801", "11401", "11702", "11702", "118", "119", "120", "121", "12304", "12305"), ]

```


### i) Household purchases (2016 to 2020)

```{r Allocate SpeciesType to food groups}
# Identify unique major food groups descriptions 
unique(df_DEFRA_household_201620$codeDESCRIPTION)

# Major food codes according to FAO groupings
df_DEFRA_household_201620$SpeciesType <- "Unknown"
unique(df_DEFRA_household_201620$SpeciesType)

df_DEFRA_household_201620$SpeciesType[df_DEFRA_household_201620$codecat %in% c("10201", "10202", "11401")] <- "Demersal"
df_DEFRA_household_201620$SpeciesType[df_DEFRA_household_201620$codecat %in% c("11702", "11703")] <- "Shellfish"
df_DEFRA_household_201620$SpeciesType[df_DEFRA_household_201620$codecat %in% c("10601", "10602", "10701", "10702", "10801", "119")]<-"Pelagic"

```

```{r Allocate SACN i.e. lean or oily}
df_DEFRA_household_201620$SACN[df_DEFRA_household_201620$SpeciesType %in% 
                                 c("Pelagic")] <- "Oily"
df_DEFRA_household_201620$SACN[df_DEFRA_household_201620$SpeciesType %in%
                                 c("Demersal")]<-"Lean"
df_DEFRA_household_201620$SACN[df_DEFRA_household_201620$SpeciesType %in%
                                 c("Unknown")]<-"Unknown"
df_DEFRA_household_201620$SACN[df_DEFRA_household_201620$SpeciesType %in%
                                 c("Shellfish")]<-"Shellfish"

```


```{r Calculate grams available per capita}
# Rename "estimate column"
names(df_DEFRA_household_201620)[names(df_DEFRA_household_201620)=="estimate"] <- "ValueGramsCapitaWeek"

```

```{r Select desired variables from data set}
# Rename columns
names(df_DEFRA_household_201620)[names(df_DEFRA_household_201620)=="codeDESCRIPTION"] <- "Species"

# Create new data frame with desired variables
dfc_DEFRA_household_201620 <- df_DEFRA_household_201620[c("SACN", "SpeciesType", "Species", "Year", "ValueGramsCapitaWeek")]

```


### ii) Eating out (2016 to 2020)

```{r Allocate SpeciesType to food groups}
# Identify unique major food groups descriptions 
unique(df_DEFRA_eaten_out_201620$codeDESCRIPTION)

# Major food codes according to FAO groupings
df_DEFRA_eaten_out_201620$SpeciesType <- "Unknown"
unique(df_DEFRA_eaten_out_201620$codecat)

df_DEFRA_eaten_out_201620$SpeciesType[df_DEFRA_eaten_out_201620$codecat %in% c("eo120102", "eo120101")] <- "Demersal"
df_DEFRA_eaten_out_201620$SpeciesType[df_DEFRA_eaten_out_201620$codecat %in% c("eo1203")] <- "Shellfish"
df_DEFRA_eaten_out_201620$SpeciesType[df_DEFRA_eaten_out_201620$codecat %in% c("eo1204", "eo120202")] <- "Pelagic"

### Trout, tuna and salmon aggregated together 

```


```{r Allocate SACN i.e. lean or oily}
# Identify unique SpeciesTypes
unique(df_DEFRA_eaten_out_201620$SpeciesType)

# State whether fish are lean or oily
df_DEFRA_eaten_out_201620$SACN[df_DEFRA_eaten_out_201620$SpeciesType %in% c("Pelagic")] <- "Oily"
df_DEFRA_eaten_out_201620$SACN[df_DEFRA_eaten_out_201620$SpeciesType %in% c("Demersal")] <- "Lean"
df_DEFRA_eaten_out_201620$SACN[df_DEFRA_eaten_out_201620$SpeciesType %in% c("Unknown")] <- "Unknown"
df_DEFRA_eaten_out_201620$SACN[df_DEFRA_eaten_out_201620$SpeciesType %in% c("Shellfish")] <- "Shellfish"

# Note
df_DEFRA_eaten_out_201620$SACN[df_DEFRA_eaten_out_201620$codecat %in% c("eo120201")] <- "Oily" # note: Trout, tuna and salmon aggregated together thus can allocate via SpeciesType

```


```{r Calculate grams available per capita}
# Raw values are provided in grams per capita so rename column
names(df_DEFRA_eaten_out_201620)[names(df_DEFRA_eaten_out_201620)=="estimate"] <- "ValueGramsCapitaWeek"

```


```{r Select desired variables from data set}
# Rename column
names(df_DEFRA_eaten_out_201620)[names(df_DEFRA_eaten_out_201620)=="codeDESCRIPTION"] <- "Species"

# Create new data frame with desired variables
colnames(df_DEFRA_eaten_out_201620)
dfc_DEFRA_eatenOut_201620 <- df_DEFRA_eaten_out_201620[c("SACN", "SpeciesType", "Species", "Year", "ValueGramsCapitaWeek")]

```


### iii) Household purchases (2009 to 2015)

```{r Load data (2009 to 2015)}
# read DEFRA household data set (.csv)
df_DEFRA_household_data <- vroom(file=paste(filepath,"DEFRA_UK_household-200915.csv", sep="/"))

# Check for missing values
md.pattern(df_DEFRA_household_data, rotate.names = TRUE)
# missinf values detected but outside of the desired timeframe

# Select only years 2009 to 2015
df_DEFRA_household_200915 <- subset(df_DEFRA_household_data, 
                           select = c("Code", "Food Category", "Food Group",
                                      "Major Food Code", "Minor Food Code",
                                      "Units", "2009", "2010", "2011",
                                      "2012", "2013", "2014", "2015")) 
# Check for missing values
md.pattern(df_DEFRA_household_200915, rotate.names = TRUE)

```


```{r Allocate SpeciesType to food groups}
# Identify unique major food groups
unique(df_DEFRA_household_200915$`Major Food Code`)

# Major food codes according to FAO groupings
df_DEFRA_household_200915$SpeciesType <- "Unknown"
df_DEFRA_household_200915$SpeciesType[df_DEFRA_household_200915$`Major Food Code` == "White fish, fresh, chilled or frozen"] <- "Demersal"
df_DEFRA_household_200915$SpeciesType[df_DEFRA_household_200915$`Major Food Code` == "White fish, dried, salted or smoked"] <- "Demersal"
df_DEFRA_household_200915$SpeciesType[df_DEFRA_household_200915$`Major Food Code` == "Shellfish"] <- "Shellfish"
df_DEFRA_household_200915$SpeciesType[df_DEFRA_household_200915$`Major Food Code` == "Herrings and other blue fish, fresh, chilled or frozen"] <-"Pelagic"
df_DEFRA_household_200915$SpeciesType[df_DEFRA_household_200915$`Major Food Code` == "Salmon, fresh, chilled or frozen"] <- "Pelagic"
df_DEFRA_household_200915$SpeciesType[df_DEFRA_household_200915$`Major Food Code` == "Blue fish, dried or salted or smoked"] <- "Pelagic"
df_DEFRA_household_200915$SpeciesType[df_DEFRA_household_200915$`Major Food Code` == "Salmon, tinned"] <- "Pelagic"

# Remove rows where sum is provided 
df_DEFRA_household_200915_clean<-df_DEFRA_household_200915[!grepl("Sum", 
                                         df_DEFRA_household_200915$`Minor Food Code`),]

```


```{r Allocate SACN i.e. lean or oily}
df_DEFRA_household_200915_clean$SACN[df_DEFRA_household_200915_clean$SpeciesType %in% c("Pelagic")] <- "Oily"
df_DEFRA_household_200915_clean$SACN[df_DEFRA_household_200915_clean$SpeciesType %in% c("Demersal")]<-"Lean"
df_DEFRA_household_200915_clean$SACN[df_DEFRA_household_200915_clean$SpeciesType %in% c("Unknown")]<-"Unknown"
df_DEFRA_household_200915_clean$SACN[df_DEFRA_household_200915_clean$SpeciesType %in% c("Shellfish")]<-"Shellfish"

```


```{r Transform and clean data set}
# Subset data frame to include variables we want before changing structure
df_DEFRA_household_200915_main<-subset(df_DEFRA_household_200915_clean, select= c("Minor Food Code", "Units", "2009", "2010", "2011", "2012", "2013", "2014","2015", "SpeciesType", "SACN"))

# Change structure of the data set 
df_DEFRA_household_200915_long<-melt(df_DEFRA_household_200915_main, id.vars = c("Minor Food Code", "Units", "SpeciesType", "SACN"), variable.name="Year", value.name="Value")

# Rename columns for consistency 
colnames(df_DEFRA_household_200915_long)
names(df_DEFRA_household_200915_long)[names(df_DEFRA_household_200915_long)=="Minor Food Code"] <- "Species"

```


```{r Calculate grams available per capita}
# Raw values are provided in grams per capita so rename column
names(df_DEFRA_household_200915_long)[names(df_DEFRA_household_200915_long)=="Value"] <- "ValueGramsCapitaWeek"

```


```{r Select desired variables from data set}
# Rename columns
names(df_DEFRA_household_200915_long)[names(df_DEFRA_household_200915_long)=="codeDESCRIPTION"] <- "Species"

# Create new data frame with desired variables
colnames(df_DEFRA_household_200915_long)
dfc_DEFRA_household_200915 <- df_DEFRA_household_200915_long[c("SACN", "SpeciesType", "Species", "Year", "ValueGramsCapitaWeek")]

```


```{r Combine DEFRA household data sets (2009 to 2015 and 2016 to 2020)}
# Combine household data (2009 to 2015 and 2016 to 2020)
colnames(dfc_DEFRA_household_200915)
str(dfc_DEFRA_household_200915)
unique(dfc_DEFRA_household_200915$Year)

colnames(dfc_DEFRA_household_201620)
str(dfc_DEFRA_household_201620)
unique(dfc_DEFRA_household_201620$Year)


# Ensure "Year" column is as a factor value
dfc_DEFRA_household_200915$Year<-as.factor(dfc_DEFRA_household_200915$Year)
dfc_DEFRA_household_201620$Year<-as.factor(dfc_DEFRA_household_201620$Year)

# Row bind data frames (2009 to 2020)
dfc_DEFRA_household_200920 <- rbind(dfc_DEFRA_household_200915, dfc_DEFRA_household_201620)

str(dfc_DEFRA_household_200920)
unique(dfc_DEFRA_household_200920$Year)

```


```{r Add units and data source}
# Create "cleaned" data frame, defining units and data source 
dfc_DEFRA_household_200920$DataSupplier <- "DEFRA"
dfc_DEFRA_household_200920$DataSet <- "Household"
dfc_DEFRA_household_200920$Commodity <- "Purchase"
dfc_DEFRA_household_200920$Flag <- "ConV2"

```


```{r Save cleaned data}
# Change file path to "ProcessedData"
filepath <- paste(data_dir,"ProcessedData", sep="")
write.csv(dfc_DEFRA_household_200920, file =
            paste(filepath,"HouseholdPurchases_DEFRA_clean.csv", sep=""))

```


### iv) Eating out (2009 to 2015)

```{r Load data (2009 to 2015)}
# read DEFRA eaten out data set
df_DEFRA_eaten_out_data <- vroom(file=paste(filepath,"DEFRA_UK_eatenOut-200915.csv", sep="/"))

# Check for missing values
md.pattern(df_DEFRA_eaten_out_data, rotate.names = T)
# Missing values detected but not within time frame of interest

# Select only years 2009 to 2015
df_DEFRA_EO_200915 <- subset(df_DEFRA_eaten_out_data, 
                           select = c("Code", "EO Food Group", 
                                      "EO Major code", "EO Minor code",
                                      "Units", "2009", "2010", "2011",
                                      "2012", "2013","2014", "2015"))

```


```{r Allocate SpeciesType to food group}
# Identify unique major food groups
unique(df_DEFRA_EO_200915$`EO Major code`)

# Major food codes according to FAO groupings.
df_DEFRA_EO_200915$SpeciesType <- "Unknown"
df_DEFRA_EO_200915$SpeciesType[df_DEFRA_EO_200915$`EO Major code` == "White fish"] <- "Demersal"
df_DEFRA_EO_200915$SpeciesType[df_DEFRA_EO_200915$`EO Major code` == "Shellfish - without sauce or dressing (e.g. prawns, shrimps, oysters, crab)"] <- "Shellfish"
df_DEFRA_EO_200915$SpeciesType[df_DEFRA_EO_200915$`EO Major code` == "Fatty fish"] <- "Pelagic"
df_DEFRA_EO_200915$SpeciesType[df_DEFRA_EO_200915$`EO Major code` == "Kippers and other smoked fish (e.g. smoked salmon)"] <- "Pelagic"

# Remove rows where sum is provided 
df_DEFRA_EO_200915_clean<-df_DEFRA_EO_200915[!grepl("Sum", 
                                         df_DEFRA_EO_200915$`EO Minor code`),]

```


```{r Allocate SACN i.e. lean or oily}
# Identify uniqye SpeciesType in the data set
unique(df_DEFRA_EO_200915_clean$SpeciesType)

# Allocate oily/lean/shellfish
df_DEFRA_EO_200915_clean$SACN[df_DEFRA_EO_200915_clean$SpeciesType %in% c("Pelagic")] <- "Oily"
df_DEFRA_EO_200915_clean$SACN[df_DEFRA_EO_200915_clean$SpeciesType %in% c("Demersal")]<-"Lean"
df_DEFRA_EO_200915_clean$SACN[df_DEFRA_EO_200915_clean$SpeciesType %in% c("Unknown")]<-"Unknown"
df_DEFRA_EO_200915_clean$SACN[df_DEFRA_EO_200915_clean$SpeciesType %in% c("Shellfish")]<-"Shellfish"

```


```{r Transform and clean data set}
# Subset data frame to include variables we want 
df_DEFRA_EO_200915_main<-subset(df_DEFRA_EO_200915_clean, select= c("EO Minor code", "Units", "2009", "2010", "2011", "2012", "2013", "2014", "2015", "SpeciesType", "SACN"))

# Change structure of the data set 
df_DEFRA_eatenOut_200915_long<-melt(df_DEFRA_EO_200915_main, id.vars = c("EO Minor code", "Units", "SpeciesType", "SACN"), variable.name="Year", value.name="Value")
```


```{r Calculate grams available per capita}
# Change column variable name to include units. Raw data values provided in g/capita/week
names(df_DEFRA_eatenOut_200915_long)[names(df_DEFRA_eatenOut_200915_long)=="Value"] <- "ValueGramsCapitaWeek"

```


```{r Select desired variables from data set}
# Rename columns
colnames(df_DEFRA_eatenOut_200915_long)
names(df_DEFRA_eatenOut_200915_long)[names(df_DEFRA_eatenOut_200915_long)=="EO Minor code"] <- "Species"

# Create new data frame with desired variables
dfc_DEFRA_eatenOut_200915 <- df_DEFRA_eatenOut_200915_long[c("SACN", "SpeciesType", "Species", "Year", "ValueGramsCapitaWeek")]

```


```{r Combine DEFRA eating out data (2009 to 2015 and 2016 to 2020)}
# Convert "Year" to a factor value
str(dfc_DEFRA_eatenOut_201620)

dfc_DEFRA_eatenOut_201620$Year <- as.factor(dfc_DEFRA_eatenOut_201620$Year)

# Combine DEFRA eating out data between 2009 to 2015 and 2015 to 2020
dfc_DEFRA_eatenOut_200920 <- rbind(dfc_DEFRA_eatenOut_200915, dfc_DEFRA_eatenOut_201620)

```


```{r Add units and data source}
# Create "cleaned" data frame, defining units and data source 
dfc_DEFRA_eatenOut_200920$DataSupplier <- "DEFRA"
dfc_DEFRA_eatenOut_200920$DataSet <- "eatenOut"
dfc_DEFRA_eatenOut_200920$Commodity <- "Purchase"
dfc_DEFRA_eatenOut_200920$Flag <- "ConV1, ConV3"

```


```{r Save cleaned data}
write.csv(dfc_DEFRA_eatenOut_200920, file =
            paste(filepath,"EatenOutPurchases_DEFRA_clean.csv", sep=""))

```



## Nutrient composition data (Public Health England)

Next we downloaded seafood nutrient composition data from Public Health England (PHE). This data includes the nutrient content (macro and micro nutrients) of commonly consumed fish in the UK. Composition analyses were conducted on raw and cooked fish samples.

Composition analyses were conducted in 2004. Temporal changes in nutrient composition is unlikely.


```{r Load data}
# Read in .csv file 
df_PHE_composition_data <- vroom(file=paste(filepath,"PHE_SeafoodNutrientContent-edit.csv", sep="/"), na = c("N/A", "Tr")) # Both N/A and Tr different 
# Note that .csv file is saved as a UT8 (otherwise an error appears. Check with Bernhard the differences in the file types)

# Check for missing values
md.pattern(df_PHE_composition_data, rotate.names = TRUE)

# Identify column names
colnames(df_PHE_composition_data)

```


```{r Rename selected individual species and provide species types}
# Duplicate species name column (duplicated column entries to be duplicated)
df_PHE_composition_data$SpeciesNEW <- df_PHE_composition_data$SampleDescription

# Merge data frame with MCS data frame by species name
df_PHE_composition_data <- df_PHE_composition_data %>% inner_join(., MCS_translations, by = c('SpeciesNEW' = 'MCS (DO NOT TOUCH)'))

```

Cleaned dataset does not consider the status of the sample e.g. raw/ cooked. Need to create a new column with such infomation. 

```{r Add additional variables to data set}
# Create "cleaned" data frame, defining units and data source 
df_PHE_composition_data$DataSupplier <- "PHE"
df_PHE_composition_data$DataSet <- "NutrientAnalysis"
df_PHE_composition_data$Commodity <- "Composition"
df_PHE_composition_data$Flag <- "ConV1"

```


```{r Select desired variables}
# Identify columns in data frame
colnames(df_PHE_composition_data)

# Select desired columns for cleaned data set
df_nutcomp_PHE_cleaned <- df_PHE_composition_data[c("RevisedMCS", "EUMOFA_ST", "FAO_ST", "ISSCAAP_ST", "Golden_2021_ST","Bianchi_2022_ST", "WWF_2022_ST", "SACN", "DataSupplier", "DataSet", "Commodity", "Flag", "Calcium-mg100g", "Iron-mg100g", "Selenium-µg100g", "Zinc-mg100g", "TotalPUFA-g100g" )]

```

Calculate nutrient density score of each seafood product. Nutrient density defines as the combined contribution of a 100g portion for recommended daily intakes of all five nutrients based on nutrient reference point for adults.

```{r Calculate nutrient density score of each fish species}
# Create new column calculating nutrient density score

```


```{r Save cleaned data}
write.csv(df_nutcomp_PHE_cleaned, file =
            paste(filepath,"NutrientComposition_PHE_clean.csv", sep=""))

```




## END

######## Code not needed 

### B) Aquaculture production (Marine Scotland and DEFRA)

```{r Load data (Scotland: Marine Scotland Science)}
# a) Aquaculture production in Scotland
# i) salmon, ii) rainbow trout, iii) shellfish
df_MSS_salmon_data <- vroom(file=paste(filepath,"MSS_Scotland_aquacultureSalmon.csv", sep="/"), na = c(".."))

df_MSS_rainbowTrout_data <- vroom(file=paste(filepath,"MSS_Scotland_aquacultureRainbowTrout.csv", sep="/"), na = c(".."))

df_MSS_shellfish_data <- vroom(file=paste(filepath,"MSS_Scotland_aquacultureShellfish.csv", sep="/"), na = c(".."))

```

```{r Subset data (Scotland: Marine Scotland Science)}
# Determine products are for human consumption 
df_MSS_shellfish_HC <- subset(df_MSS_shellfish_data, Production == "For the table")

# Subset data between 2009 and 2021 
df_MSS_salmon <- subset(df_MSS_salmon_data, Year > 2008, 
                        select= c(Year, Tonnes)) 

df_MSS_rainbowTrout <- subset(df_MSS_rainbowTrout_data, Year > 2008, 
                        select= c("Year", "Total Tonnes"))

df_MSS_shellfish <- subset(df_MSS_shellfish_HC, Year > 2008)

# State species, unit and data source
df_MSS_salmon$Units <- "tonnes"
df_MSS_salmon$Species <- "Salmon"
df_MSS_salmon$SpeciesType <- "Diadromous"

df_MSS_rainbowTrout$Units <- "tonnes"
df_MSS_rainbowTrout$Species <- "RainbowTrout"
df_MSS_rainbowTrout$SpeciesType <- "Freshwater"

df_MSS_shellfish$Units <- "Varied"
df_MSS_shellfish$Species <- "Varied"
df_MSS_rainbowTrout$SpeciesType <- "Shellfish"

```

```{r Combine Scottish aqacualture production data}

#df_MMS_production<- merge(df_MSS_salmon, df_MSS_rainbowTrout)

```


```{r Add units and data source (Scotland: Marine Scotland Science)}
# Create "cleaned" data frame, defining units and data source 

#df_clean_UK_landings_09_20$Units <- "tonnes or numbers"
df_MMS_production$DataSupplier <- "Marine_Scotland"
df_MMS_production$DataSet <- "Scottish-aquaculture"
df_MMS_production$Commodity <- "Production"
df_MMS_production$TempRes <- "Annual"
```


### B) Aquaculture production (CEFAS)

```{r Load data (2013 to 2020)}
# UK aquaculture production 
df_CEFAS_UK_aquaculture_data <- vroom(file=paste(filepath,"CEFAS_UK_aquaculture-201320-edit.csv", sep="/"))

# Rename column
names(df_CEFAS_UK_aquaculture_data)[names(df_CEFAS_UK_aquaculture_data)=="Tonnes"] <- "Units"
```

```{r Transform and clean data set (2013 to 2010)}
# Change structure of the data set 
df_UK_aquaculture_201320_long<-melt(df_CEFAS_UK_aquaculture_data, id.vars = c("Type", "Species", "Units"), variable.name="Year", value.name="Value")

```

```{r Allocating fish species corresonding species type (2013 to 2010)}
# Allocate fish species group type 

```

```{r Load data (2009 to 2012)}
# UK aquaculture production 
df_CEFAS_UK_aqaculture_data <- vroom(file=paste(filepath,".csv", sep="/"))

```


```{r Transform and clean data set (2009 to 2012)}
# Change structure of the data set 
df_UK_aquaculture_201320_long<-melt(df_CEFAS_UK_aqaculture_data, id.vars = c("Type", "Species", "Units"), variable.name="Year", value.name="Value")

```

```{r Allocating fish species corresonding species type (2009 to 2012)}
# Allocate fish species group type 

```


```{r Add units and data source (2013 to 2020)}
# Create "cleaned" data frame, defining units and data source 
df_UK_aquaculture_201320_long$DataSupplier <- "CEFAS"
df_UK_aquaculture_201320_long$DataSet <- "Aquaculture"
df_UK_aquaculture_201320_long$Commodity <- "Production"
df_UK_aquaculture_201320_long$TempRes <- "Annual"
```

```{r Save cleaned data (CEFAS 2013 to 2020)}

```

####
```{r}
# Rename column
names(df_UK_landings_11_20_data)[names(df_UK_landings_11_20_data)=="Species_type"] <- "SpeciesType"
```

```{r Load data (2011 to 2020)}
# i) Landings into the UK by UK vessels: 2011 to 2020 (a)

# Read in .csv file
df_UK_landings_11_20_data <- vroom(file=paste(filepath,"NS_UK_landings-201120-edit.csv", sep="/"), na = c(".."))
```


```{r Provide greater granularity in "SpeciesType" column}
# Provide greater granularity to shellfish species
UK_shellfishLandings_11_20 <- subset(df_UK_landings_11_20_data, SpeciesType == "Shellfish") 

# Default "SpeciesType" to "mollusc"
UK_shellfishLandings_11_20$SpeciesType <- "Mollusc"

# Allocate crustaceans 
UK_shellfishLandings_11_20$SpeciesType[UK_shellfishLandings_11_20$Species %in% c("Crabs", "Lobsters", "Nephrops", "Shrimps and Prawns")] <- "Crustacean"

# Allocate cephalopods
UK_shellfishLandings_11_20$SpeciesType[UK_shellfishLandings_11_20$Species %in% c("Cuttlefish", "Squid")] <- "Cephalopod"

# Allocate other shellfish
UK_shellfishLandings_11_20$SpeciesType[UK_shellfishLandings_11_20$Species %in% "Other Shellfish"] <- "Other shellfish"

# Combine with data set 
UK_finfishLandings_11_20 <- df_UK_landings_11_20_data[(df_UK_landings_11_20_data$SpeciesType == "Pelagic") | (df_UK_landings_11_20_data$SpeciesType == "Demersal"),]

df_UK_landings_11_20 <- rbind(UK_finfishLandings_11_20, UK_shellfishLandings_11_20)

```

```{r Transform and clean data set (2011 to 2020)}
# Change structure of the data set 
df_UK_landings_201120_long<-melt(df_UK_landings_11_20, id.vars = c("SpeciesType","Species"), variable.name="Year", value.name="Value")

```


## UK population data (Office of National Statistics)

Download UK population data- need to convert per g/capita data to g/total UK total population.

```{r Load data}
# Read in UK population data 
df_NS_UK_pop_data <- vroom(file=paste(filepath,"NS_UK_population.csv", sep="/"))
```

```{r Subset data}
# Extract UK population data between 2009 onwards
df_UK_pop <- subset(df_NS_UK_pop_data, 
                      CDID > 2008, select= c(CDID, UKPOP))

```

```{r Add units and data source}
# Create "cleaned" data frame, defining units and data source 
df_clean_UK_pop <- df_UK_pop
df_clean_UK_pop$Units <- "UK-total-pop"
df_clean_UK_pop$Data_source <- "National-statistics"
```


```{r Save cleaned data}
# Change file path to "ProcessedData". Can also use setwd()
filepath <- paste(data_dir,"ProcessedData/csv-files", sep="")
filepath <- paste(data_dir,"ProcessedData", sep="")
write.csv(df_clean_UK_pop, file = paste(filepath,"PopSize_NS.csv", sep=""))

```

```{r}
# Changes species one at a time 
str_replace_all(df_UK_landings_09_20$Species, "Blue Whiting", "Blue whiting")
str_replace_all(df_UK_landings_09_20$Species, "Cockles", "Cockle")
str_replace_all(df_UK_landings_09_20$Species, "Crabs", "Crab")
#str_replace_all(df_UK_landings_09_20$Species, "Horse Mackerel", "Horse mackerel")
#str_replace_all(df_UK_landings_09_20$Species, "Lobsters", "Lobster")
#str_replace_all(df_UK_landings_09_20$Species, "Monks or Anglers", "Monk")
#str_replace_all(df_UK_landings_09_20$Species, "Mussels", "Mussel")
#str_replace_all(df_UK_landings_09_20$Species, "Pollack (Lythe)", "Pollock")
#str_replace_all(df_UK_landings_09_20$Species, "Sand Eels", "Sandeel")
#str_replace_all(df_UK_landings_09_20$Species, "Sardines", "Sardine")
#str_replace_all(df_UK_landings_09_20$Species, "Shrimps and Prawns", "Shrimp")
#str_replace_all(df_UK_landings_09_20$Species, "Skates and Rays", "Skate")
#str_replace_all(df_UK_landings_09_20$Species, "Whelks", "Whelk")

#df_UK_landings_09_20$Species == "Pollack (Lythe)"
```


```{r Calculate grams available per capita}

# Add in annual UK population data in a separate column. Duplicate population data for each year. Number of replicates dictated by number of rows in the short and narrow data set.
#df_UK_landings_200920$PopSize<- rep(c(df_clean_UK_pop$UKPOP), each = nrow(df_UK_landings_08_12_data))

# Convert grams/year to g/capita/year
#df_UK_landings_200920$ValueGramsCapitaYear<- #df_UK_landings_200920$ValueGrams/df_UK_landings_200920$PopSize

# Convert grams/year to g/capita/week
#df_UK_landings_200920$ValueGramsCapitaWeek<- df_UK_landings_200920$ValueGramsCapitaYear/52

```


```{r Calculate grams available per capita}
# Add in annual UK population data in a separate column. Duplicate population data for each year. Number of replicates dictated by number of rows in the short and narrow data set.
#df_EUROSTAT_UKFarmed$PopSize<- rep(c(df_clean_UK_pop$UKPOP), each = #nrow(df_EUROSTAT_UK_aquaculture_data))

# Convert raw value from tonnes to grams 
#df_EUROSTAT_UKFarmed$ValueGrams<- (df_EUROSTAT_UKFarmed$CFValue*1000000)

# Convert grams/year to g/capita/year
#df_EUROSTAT_UKFarmed$ValueGramsCapitaYear<- #df_EUROSTAT_UKFarmed$ValueGrams/df_EUROSTAT_UKFarmed$PopSize

# Convert grams/year to g/capita/week
#df_EUROSTAT_UKFarmed$ValueGramsCapitaWeek<- df_EUROSTAT_UKFarmed$ValueGramsCapitaYear/52

```

```{r}
# Number of seafood entries differ per year so need to create a pipe
df_trade_HMRC_data$popSize[df_trade_HMRC_data$Year %in% "2020"] <- 67081000
df_trade_HMRC_data$popSize[df_trade_HMRC_data$Year %in% "2019"] <- 66796800
df_trade_HMRC_data$popSize[df_trade_HMRC_data$Year %in% "2018"] <- 66435600
df_trade_HMRC_data$popSize[df_trade_HMRC_data$Year %in% "2017"] <- 66040200
df_trade_HMRC_data$popSize[df_trade_HMRC_data$Year %in% "2016"] <- 65648100
df_trade_HMRC_data$popSize[df_trade_HMRC_data$Year %in% "2015"] <- 65110000
df_trade_HMRC_data$popSize[df_trade_HMRC_data$Year %in% "2014"] <- 64596800
df_trade_HMRC_data$popSize[df_trade_HMRC_data$Year %in% "2013"] <- 64105700
df_trade_HMRC_data$popSize[df_trade_HMRC_data$Year %in% "2012"] <- 63705000
df_trade_HMRC_data$popSize[df_trade_HMRC_data$Year %in% "2011"] <- 63285100
df_trade_HMRC_data$popSize[df_trade_HMRC_data$Year %in% "2010"] <- 62759500
df_trade_HMRC_data$popSize[df_trade_HMRC_data$Year %in% "2009"] <- 62260500

# Make popsize column numerical value
df_trade_HMRC_data$popSize <- as.numeric(df_trade_HMRC_data$popSize)

# Calculate grams/capita/year
df_trade_HMRC_data$ValueGramsCapitaYear <- df_trade_HMRC_data$ValueGrams/df_trade_HMRC_data$popSize

# Calculate grams/capita/week
df_trade_HMRC_data$ValueGramsCapitaWeek <- df_trade_HMRC_data$ValueGramsCapitaYear/52



# Replace all occurrences of a certain pattern i.e. the species names
df_UK_landings_09_20$SpeciesNEW <- str_replace_all(df_UK_landings_09_20$SpeciesNEW, 
                   c("Blue Whiting" = "Blue whiting", "Cockles" = "Cockle", 
                     "Crabs" = "Crab", "Horse Mackerel" = "Horse mackerel",
                     "Lobsters" = "Lobster", "Lemon sole" = "Sole", 
                     "Monks or Anglers" = "Monks", "Mussels" = "Mussel", 
                     "Pollack (Lythe)" = "Pollock", "Sand Eels" = "Sandeel",
                     "Sardines" = "Sardine", "Shrimps and Prawns" = "Shrimp", 
                     "Skates and Rays" = "Ray", "Whelks" = "Whelk", 
                     "Bass" = "Seabass", "Other Demersal (b)" = "Other demersal",
                     "Other Pelagic" = "Other pelagics", "Scallops" = "Scallop",
                     "Other Shellfish" = "Other shellfish"))


# str_replace_all() does not work with brackets
```


```{r}
# Identify unique SpeciesType entries
#unique(df_UK_landings_200920$SpeciesType)

# Create a new column in the data frame and allocate oily/lean/shellfish classification
#df_UK_landings_200920$SACN[df_UK_landings_200920$SpeciesType == "Crustacean"] <- "Shellfish"
#df_UK_landings_200920$SACN[df_UK_landings_200920$SpeciesType == "Mollusc"] <- "Shellfish"
#df_UK_landings_200920$SACN[df_UK_landings_200920$SpeciesType == "Cephalopod"] <- "Shellfish"
#df_UK_landings_200920$SACN[df_UK_landings_200920$SpeciesType == "OtherShellfish"] <- #"Shellfish"
#df_UK_landings_200920$SACN[df_UK_landings_200920$SpeciesType == "Demersal"] <- "Lean"
#df_UK_landings_200920$SACN[df_UK_landings_200920$SpeciesType == "Pelagic"] <- "Oily"

# No freshwater fish? 
#df_UK_landings_200920$SACN[df_UK_landings_200920$SpeciesType == "Freshwater"] <- "Oily"

```


```{r Allocate SACN values}
# Identify unique SpeciesTypes in the data set
unique(df_trade_HMRC_data$SpeciesType)

# Create a new column for SACN values
df_trade_HMRC_data$SACN[df_trade_HMRC_data$SpeciesType == "Crustacean"] <- "Shellfish"
df_trade_HMRC_data$SACN[df_trade_HMRC_data$SpeciesType == "Mollusc"] <- "Shellfish"
df_trade_HMRC_data$SACN[df_trade_HMRC_data$SpeciesType == "Cephalopod"] <- "Shellfish"
df_trade_HMRC_data$SACN[df_trade_HMRC_data$SpeciesType == "Freshwater"] <- "Oily"
df_trade_HMRC_data$SACN[df_trade_HMRC_data$SpeciesType == "Demersal"] <- "Lean"
df_trade_HMRC_data$SACN[df_trade_HMRC_data$SpeciesType == "Pelagic"] <- "Oily"

# Also state unknowns
df_trade_HMRC_data$SACN[df_trade_HMRC_data$SpeciesType == "OtherFishProducts"] <- "NoData"
df_trade_HMRC_data$SACN[df_trade_HMRC_data$SpeciesType == "Marine fish"] <- "NoData"
df_trade_HMRC_data$SACN[df_trade_HMRC_data$SpeciesType == "Invertebrate"] <- "NoData"


# Replace all occurrences of a certain pattern i.e. the species names
df_trade_HMRC_data$SpeciesNEW <- 
  str_replace_all(df_trade_HMRC_data$SpeciesNEW, 
                   c("Tuna, bluefin" = "Tuna", 
                   "Caviar, livers and roes" = "Other fishProducts"))

# Identify unique species names in trade data  
unique(df_trade_HMRC_data$SpeciesNEW)

```


## Total UK fish consumption (NDNS) OLD

Total fish consumption for the UK by population cohort was downloaded between 2009 and 2019. Below is how the data were cleaned but unlikely to be added to the data base due to a lack of granularity. Data set is reported for every financial year (?).

```{r Load data}
# Read in .csv file
df_NDNS_UK_data <- vroom(file=paste
                      (filepath,"NDNS_UK_consumption-edit.csv", sep="/"))

```


```{r Aggregate cohort information}
# Identify unique age cohorts
unique(df_NDNS_UK_data$`Age category`)

# Create a pipe aggregating sub age cohort totals, irrespective of sex 
NDNS_totals <- df_NDNS_UK_data[df_NDNS_UK_data$`Age category` %in% c("Children 1.5-3 years", "Children 4-10 years", "Children 11-18 years", "Adults 19-64 years", "Adults 65 years and over"), ]

# Sum all consumption data for all years 
df_NDNSSummed <- as.data.frame(apply(NDNS_totals[3:7], 2, sum))

df_NDNSSummed <- as.data.frame(apply(NDNS_totals[c("(2008/09 - 2009/10)", 
"(2010/11 - 2011/12)", "(2012/13 - 2013/14)", "(2014/15-2015/16)",  
"(2016/17-2018/19)")], 2, sum))

df_NDNSSummed <- as.data.frame(colSums(NDNS_totals[c("(2008/09 - 2009/10)", 
"(2010/11 - 2011/12)", "(2012/13 - 2013/14)", "(2014/15-2015/16)",  
"(2016/17-2018/19)")]))

# Rename first column 
df_NDNSSummed<- colnames("Year", "Value")

####

NDNSYears <- c("(2008/09 - 2009/10)", 
"(2010/11 - 2011/12)", "(2012/13 - 2013/14)", "(2014/15-2015/16)",  
"(2016/17-2018/19)")

SumnedNDNS <- as.data.frame(colSums(NDNS_totals[c("(2008/09 - 2009/10)", 
"(2010/11 - 2011/12)", "(2012/13 - 2013/14)", "(2014/15-2015/16)",  
"(2016/17-2018/19)")]))

df_NDNSSummed<- data.frame(NDNS_colnams, SumnedNDNS)


```

```{r Add units and data source}
# Create "cleaned" data frame, defining units and data source 
df_clean_NDNS$DataSupplier <- "NDNS"
df_clean_NDNS$DataSet <- "NDNS"
df_clean_NDNS$Commodity <- "Consumption"
df_clean_NDNS$TempRes <- ""

# SpeciesType, SACN and Species unknown
df_clean_NDNS$SpeciesType <- "Unknown"
df_clean_NDNS$SACN <- "Unknown"
df_clean_NDNS$Species <- "Unknown"
```


```{r Select desired variables for the data set}
dfc_UK_NDNS <- df_clean_NDNS[c("SACN", "SpeciesType", "Year", "ValueGramsCapitaWeek", "DataSupplier", "DataSet", "Commodity", "flag")]

```



AL to check code below to see if needed. Code provides greater granularity to SpeciesType allocation as provided by the Office of National Statistics. 

```{r Provide greater granularity in the SpeciesType column OLD}
# Provide greater granularity in the "SpeciesType" column for shellfish
UK_shellfishLandings_09_20 <- subset(df_UK_landings_09_20, SpeciesType == "Shellfish") 
# Default "SpeciesType" to "Mollusc"
UK_shellfishLandings_09_20$SpeciesType <- "Mollusc"

# Allocate crustaceans 
UK_shellfishLandings_09_20$SpeciesType[UK_shellfishLandings_09_20$SpeciesNEW %in% c("Crab", "Lobster", "Nephrops", "Shrimp")] <- "Crustacean"

# Allocate cephalopods
UK_shellfishLandings_09_20$SpeciesType[UK_shellfishLandings_09_20$SpeciesNEW %in% c("Cuttlefish", "Squid")] <- "Cephalopod"

# Allocate other shellfish
UK_shellfishLandings_09_20$SpeciesType[UK_shellfishLandings_09_20$SpeciesNEW %in% "Other Shellfish"] <- "OtherShellfish"

# Combine with data set 
# Subset finfish landings 
UK_finfishLandings_09_20 <- df_UK_landings_09_20[(df_UK_landings_09_20$SpeciesType == "Pelagic") | (df_UK_landings_09_20$SpeciesType == "Demersal"),]

# Combine shellfish and finfish landings 
df_UK_landings_200920 <- rbind(UK_finfishLandings_09_20, UK_shellfishLandings_09_20)

```
