---
title: "NatureFood_SeafoodFigures"
author: "Lofstedt, A."
date: "18-04-23"
output: powerpoint_presentation
---

### Preparation 

```{r Load libraries, include = FALSE}
# Load packages
library(tidyr) # for tidying data
library(dplyr) # base R alternative but neater
library(vroom) # for loading and transforming data
library(tidyverse) # data exploration 
library(reshape2) # for melting data frames i.e. short and wide to long and thin
library(data.table) # for fread()-function
library(mice) # md.pattern to show missing data
library(plyr) # joins multiple data frames together
library(ggplot2) # makes plots fancier
library(scales) # displays integers on the axis
library(ggthemes) # contains extra themes, scales and functions for and related to ggplot2
library(gridExtra)
library(cowplot) # Plots two figures one output

```


```{r Edit chunk settings, include = FALSE}
knitr::opts_chunk$set(
  echo = FALSE, 
  message = FALSE, 
  warninf = FALSE, 
  fig.height = 9, 
  fig.width = 16
)
```


```{r Setup default for figures, include=FALSE}
# Set default presentation of figures
theme_set(theme_bw()%+replace%
            theme(
              title = element_text(size = 26), 
              text = element_text(size = 30), 
              axis.text = element_text(size = 24, colour = "gray40"), 
              legend.key.size = unit(1.2, "cm"), 
              legend.title = element_text(size = 26), 
              legend.text = element_text(size = 24),
              panel.grid.major = element_line(colour= "gray40", linetype = "dotted", linewidth = 0.9), 
              panel.grid.minor = element_line(colour = "gray70", linetype = "dotted", linewidth = 0.7)
            ))
# instead of "size" can also use "linewidth"
```


```{r Set colour-blind palette}
# Colour-blind friendly palette with grey. Try to keep the order (http://www.cookbook-r.com/Graphs/Colors_(ggplot2)/#a-colorblind-friendly-palette)
cbPalette <- c( "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7", "#999999")
#consumption, exports, imports, production, purchases, supplies

# Colours will be allocated in commodity alphabetical order. 
```


```{r Load file path, echo=TRUE}
# Loading file path to the data into a value 
source("Data_filepath.R")# Data_filepath.R is listed in .gitignore-file. So, you will need to create that file yourself and provide your respected filepath using "data_dir <- [enter your here]"
```


```{r Specify the desired file path}
filepath <- paste(data_dir,"ProcessedData", sep = "") 
# note we are able to change this to another file e.g. saving the output
```

```{r Load population data}
# Load database 
df_UK_population <- vroom(file=paste(filepath,"UKPopulation_ONS_PreliminaryCleaned.csv", sep="/"), na = c("NA"))
```

```{r Specify the desired file path}
filepath <- paste(data_dir,"Outputs", sep = "") 
# note we are able to change this to another file e.g. saving the output
```

```{r Load the seafood database}
# Load database 
df_UK_seafoodDB <- vroom(file=paste(filepath,"Seafood-database_preliminary.csv", sep="/"), na = c("NA"))
```


## Final clean of the dataset

Remove products for non-human consumption nor are included in the dietary recommendations 

```{r Final clean}
# Check the unique species types
unique(df_UK_seafoodDB$SpeciesType)

# Remove species type products not included in dietary recommendations e.g. "Fish meal", "white fish", "fish oil", "Seaweed and other algae". Remove products for non-human consumption. Remove 2019 data
df_UK_seafoodDB_food <- subset(df_UK_seafoodDB, SpeciesType != "Fish oil" & SpeciesType != "Seaweed and other algae" & Species != "Fishmeal" & Species != "Caviar, livers and roes" & Species != "White fish" & Species != "Surimi" & Year != 2019)
```


```{r Change structure of the data frame}
# See pivot_wider()/ spread() command in tidyverse to change structure. 
# See dplyr cheat sheet.

# Simplify data set and select desired columns 
df_UK_seafoodDB_food_spec <- df_UK_seafoodDB_food[c("Species", "SpeciesType", "SACN", "Year", "Commodity", "Value" )]

# Aggregate value by commodity, year and species (production data composed of aquaculture and wild capture)
df_UK_seafoodDB_food_simp <- aggregate(Value ~ Commodity + Year + Species + SpeciesType + SACN, data = df_UK_seafoodDB_food_spec, sum)

# Make data frame wide and short
df_UK_seafoodDB_short <- df_UK_seafoodDB_food_simp %>%
  pivot_wider(names_from = Commodity, values_from = Value)

```


```{r Combine database with UK population data}
# Merge seafood database with UK population data
df_UK_seafoodDB_Pop <- merge(df_UK_seafoodDB_short, df_UK_population, by = "Year")

# Simplify data set and select desired columns 
df_UK_seafoodDB_capita <- df_UK_seafoodDB_Pop[c("Species", "SpeciesType", "SACN", "Year", "Production",
                                                   "Exports","Imports", "Consumption","Purchases",
                                                   "UKPOP")]

```


## Calculate UK seafood supplies (g/capita/week)

```{r Calculate seafood supplies (g/capita/week)}
# Convert NA to zeros (otherwise supplies cannot be calculated)
df_UK_seafoodDB_capita[is.na(df_UK_seafoodDB_capita)] <- 0

# Calculate seafood supplies
df_UK_seafoodDB_capita$Supplies <- (df_UK_seafoodDB_capita$Production + df_UK_seafoodDB_capita$Imports) - df_UK_seafoodDB_capita$Exports

# Calculate seafood supplied per capita per week
df_UK_seafoodDB_capita$Supplies_capWeek <- (df_UK_seafoodDB_capita$Supplies/ df_UK_seafoodDB_capita$UKPOP)/52

# Calculate seafood consumed per capita per week 
df_UK_seafoodDB_capita$Consumption_capWeek <- (df_UK_seafoodDB_capita$Consumption/ df_UK_seafoodDB_capita$UKPOP)/52

```




### Figure 1. Total UK seafood supplies (g/capita/week) between 2009 and 2018  

```{r Total UK seafood supplies (g/capita/week) between 2009 and 2018}
# Sum annual seafood supplies per capita. 
df_UK_supplies_tot <- aggregate(Supplies_capWeek ~ Year, data = df_UK_seafoodDB_capita, sum)

# Plot total seafood supplies (g/capita/week) between 2009 and 2018. 
ggplot(df_UK_supplies_tot, aes(Year, Supplies_capWeek)) +
  geom_bar(stat= "identity", color = "black", fill= "#009E73" ) +
  scale_y_continuous(name = "Total seafood supplies (g/capita/week)", limits = c(0,300)) +
  scale_x_continuous(name = "Year", breaks = seq(2009, 2018, 1)) +
  geom_hline(yintercept = 280, linetype = "dashed", color = "red", linewidth = 2) +
  #geom_hline(yintercept = 140, linetype = "twodash", color = "navyblue", linewidth = 2) +
  theme(legend.title = element_blank()) 

```



### Figure 1B. Total UK seafood supplies and consummed (g/capita/week) between 2009 and 2018  

```{r Total UK seafood supplied and consumed (g/capita/week) between 2009 and 2018}
# Sum annual seafood supplies per capita 
df_UK_supplies_tot <- aggregate(Supplies_capWeek ~ Year, data = df_UK_seafoodDB_capita, sum)
df_UK_consummed_tot <- aggregate(Consumption_capWeek ~ Year, data = df_UK_seafoodDB_capita, sum)

# Rename data frame variables 
names(df_UK_supplies_tot) <- c("Year", "Value")
names(df_UK_consummed_tot) <- c("Year", "Value")

# Add additional variable
df_UK_supplies_tot$Commodity <- "Supplies"
df_UK_consummed_tot$Commodity <- "Consumption"

# rbind data frames
df_UK_supplies_intake_tot<- rbind(df_UK_supplies_tot, df_UK_consummed_tot)

# Plot total seafood supplied and consumed (g/capita/week) between 2009 and 2018. 
ggplot(df_UK_supplies_intake_tot, aes(Year, Value, fill= Commodity)) +
  geom_bar(position = "dodge", stat= "identity", color = "black") +
  scale_y_continuous(name = "Amount of seafood (g/capita/week)", limits = c(0,300)) +
  scale_x_continuous(name = "Year", breaks = seq(2009, 2018, 2)) +
  geom_hline(yintercept = 280, linetype = "dashed", color = "red", linewidth = 2) +
  geom_hline(yintercept = 140, linetype = "twodash", color = "navyblue", linewidth = 2) +
  scale_fill_manual(values = c("#E69F00", "#009E73")) +
  theme(legend.title = element_blank()) 

```


### Figure 2. UK seafood supplies and consumption (g/capita/week), by species, in 2018

```{r Subset seafood database to 2018}
# Select variables 
df_UK_seafoodCapita_suppCons <- df_UK_seafoodDB_capita[c("Supplies_capWeek", "Consumption_capWeek", "Year", "Species", "SpeciesType", "SACN")]

# Sum annual seafood supplies per capita 
df_UK_seafoodCapita_suppCons_2018 <- subset(df_UK_seafoodCapita_suppCons, Year == "2018")

# Change structure of the data set 
df_UK_seafoodCapita_suppCons_2018_long <- melt(df_UK_seafoodCapita_suppCons_2018, id.vars = c("Year", "Species", "SpeciesType", "SACN"), variable.name = "Commodity", value.name = "Value")

# Aggregate data 
df_UK_suppCons_2018_sum <- aggregate(Value ~ Species + SpeciesType + SACN + Commodity, data = df_UK_seafoodCapita_suppCons_2018_long, sum)

# Omit zero readings
df_UK_suppCons_2018_sum <- na.omit(df_UK_suppCons_2018_sum)

# Plot total seafood supplied and consumed (g/capita/week) between 2009 and 2018. 
ggplot(df_UK_suppCons_2018_sum, aes(Species, Value, fill= Commodity)) +
  geom_bar(position = "dodge", stat= "identity", color = "black") +
  scale_y_continuous(name = "Amount of seafood (g/capita/week)", limits = c(0,35)) +
  scale_x_discrete(guide = guide_axis(angle = 90), name = "") +
  geom_hline(yintercept = 280, linetype = "dashed", color = "red", linewidth = 2) +
  scale_fill_manual(values = c("#009E73", "#E69F00")) +
  theme(legend.title = element_blank()) +
  theme(legend.position = "none")

```



### Figure 2B. Selected UK seafood supplied and consummed (g/capita/week), to species level in 2018.  

```{r Subset desired seafood species in 2018 only}
# Subset 2018 data only
df_UK_seafoodDB_capita_2018 <- subset(df_UK_seafoodCapita_suppCons, Year == 2018)

# Identify seafood species 
unique(df_UK_seafoodDB_capita$Species)

# Subset specific species of interest
df_UK_SOI_capita_2018 <- subset(df_UK_seafoodDB_capita_2018, Species == "Haddock" | 
                                  Species == "Mackerel" | Species == "Salmon" | Species == "Tuna" | 
                                  Species == "Whiting"| Species == "Shrimp"| Species == "Herring" |
                                  Species == "Halibut"| Species == "Plaice" | Species == "Cockle" | 
                                  Species == "Seabass" | Species == "Blue whiting" | 
                                  Species == "Trout" | Species ==  "Octopus" | Species == "Cod")

# Simplify data set and select desired columns 
df_UK_SOI_capita_2018_specific <- df_UK_SOI_capita_2018[c("Species", "SpeciesType", "SACN", "Consumption_capWeek", "Supplies_capWeek")]

# Change structure of the data set 
df_UK_SOI_capita_2018_short <- melt(df_UK_SOI_capita_2018_specific, id.vars = c("Species", "SpeciesType", "SACN"), variable.name = "Commodity", value.name = "Value")

# Plot total seafood supplied and consumed (g/capita/week) in 2018 
ggplot(df_UK_SOI_capita_2018_short, aes(Species, Value, fill= Commodity)) +
  geom_bar(position = "dodge", stat= "identity", color = "black") +
  scale_y_continuous(name = "Amount of seafood (g/capita/week)", limits = c(0,35)) +
  scale_x_discrete(guide = guide_axis(angle = 90), name = "") +
  geom_hline(yintercept = 280, linetype = "dashed", color = "red", linewidth = 2) +
  scale_fill_manual(values = c("#009E73", "#E69F00")) +
  theme(legend.title = element_blank()) +
  theme(legend.position = "none")


# Same plot but reorder species
ggplot(df_UK_SOI_capita_2018_short, aes(reorder(Species, -Value), Value, fill= Commodity)) +
  geom_bar(position = "dodge", stat= "identity", color = "black") +
  scale_y_continuous(name = "Amount of seafood (g/capita/week)", limits = c(0,35)) +
  scale_x_discrete(guide = guide_axis(angle = 90), name = "") +
  geom_hline(yintercept = 280, linetype = "dashed", color = "red", linewidth = 2) +
  scale_fill_manual(values = c("#009E73", "#E69F00")) +
  theme(legend.title = element_blank()) +
  theme(legend.position = "none")
  
```


### Figure 3. Total UK seafood produced (g/capita/week) for oily and lean fish in 2018. 

```{r Subset production data for 2018 only}
# Subset production commodity 
df_UK_seafoodProduction <- df_UK_seafoodDB_capita[c("Production", "Year", "Species", "SpeciesType", "SACN", "UKPOP")]

# Calculate UK production per capita
df_UK_seafoodProduction$ProductionCapita <- df_UK_seafoodProduction$Production / df_UK_seafoodProduction$UKPOP
df_UK_seafoodProduction$ProductionCapitaWeek <- df_UK_seafoodProduction$ProductionCapita / 7

# Subset data for 2018
df_UK_SFProductionCapita_18 <- subset(df_UK_seafoodProduction, Year == "2018")

# Sum UK seafood production, per SACN
df_UK_SFProduction_cap_2018_SACN <- aggregate(ProductionCapitaWeek ~ SACN + Year, data = df_UK_SFProductionCapita_18, sum)

# Select for lean and oily fish only
df_UK_production_2018 <- subset(df_UK_SFProduction_cap_2018_SACN, SACN == "Lean" | SACN == "Oily")

# Plot average seafood supply (g/capita/week), to SACN level (lean and oily) between 2009 and 2018. 
ggplot(df_UK_production_2018, aes(y = ProductionCapitaWeek, x = Year, fill = SACN)) +
  geom_bar(position = "stack", stat = "identity", color = "black", width = 0.5) + 
  scale_y_continuous(name = "Total seafood production (g/capita/week)") +
  scale_x_discrete(name = "") +
  scale_fill_manual(values = c("#CC79A7", "#D55E00"))+
  geom_hline(yintercept = 280, linetype = "dashed", color = "red", linewidth = 2.0)

```



### Figure 4. Nutrient supply, by species, in 2018. 

Key nutrients include calcium, iron, vitamin D, vitamin A, vitamin B12, iodine, selenium, omega-3 fatty acids and zinc
Sum nutrient produced, imported and exported to calculate supplies.


```{r Specify the desired file path}
# Select file path with raw data
filepath <- paste(data_dir,"RawData/csv-files", sep = "") 
```


```{r Load raw PHE nutrient compostion data}
# Load data
df_PHE_seafoodComp <- vroom(file=paste(filepath,"PHE_seafoodNutrientContent.csv", sep="/"), na = c("N/A"))

# Select desired variables
df_PHE_seafoodCompSimp <- df_PHE_seafoodComp[c("SampleDescription", "Protein-g100g", "TotalPUFA-g100g", "VitaminA-µg100g", "VitaminD-µg100g", "VitaminB12-µg100g", "Zinc-mg100g", "Iron-mg100g", "Calcium-mg100g", "Selenium-µg100g", "Iodine-µg100g")]

```



```{r Load raw Norwegian nutrient compostion data}
# Load data 
df_Norwegian_seafoodComp <- vroom(file=paste(filepath,"Norwegian_seafoodNutrientContent_V2.csv", sep="/"), na = c("M", ""))

# Select desired variables
df_Norwegian_seafoodCompSimp <- df_Norwegian_seafoodComp[c("FoodItem", "Protein_g", "Omega3_g", "VitaminA_RAE", "VitaminD_µg", "VitaminB12_µg", "Zinc_mg", "Iron_mg", "Calcium_mg", "Selenium_µg", "Iodine_µg")]
```



```{r Specify the desired file path}
filepath <- paste(data_dir,"Methods/SpeciesTypeClassification", sep = "") 
# note we are able to change this to another file e.g. saving the output
```


```{r Load translation document}
# Load database 
df_seafoodCompKey <- vroom(file=paste(filepath,"SeafoodComposition_translationKey.csv", sep="/"), na = c("NA"))
```


```{r Combine nutrient composition data}
# Rename variables
names(df_PHE_seafoodCompSimp) <- c("FoodItem", "Protein_g", "TotalPUFA_g", "VitaminA_µg", "VitaminD_µg", "VitaminB12_µg", "Zinc_mg", "Iron_mg", "Calcium_mg", "Selenium_µg", "Iodine_µg")

names(df_Norwegian_seafoodCompSimp) <- c("FoodItem", "Protein_g", "TotalPUFA_g", "VitaminA_µg", "VitaminD_µg", "VitaminB12_µg", "Zinc_mg", "Iron_mg", "Calcium_mg", "Selenium_µg", "Iodine_µg")

# Combine data frames
df_seafoodCompSimp <- rbind(df_PHE_seafoodCompSimp,
                              df_Norwegian_seafoodCompSimp) 
```



```{r Join data with translation document}
# Join data frame with translation document- ensures consistent spelling across data sets

# Duplicate species name column (allows us to check species have been renames correctly)
df_seafoodCompSimp$SpeciesRevised <- df_seafoodCompSimp$FoodItem  

# Join data frame with MCS data frame by species name
df_seafoodCompSimp <- df_seafoodCompSimp %>% inner_join(., df_seafoodCompKey, by = c('SpeciesRevised' = 'MCS (DO NOT TOUCH)'))
# MSC (DO NOT TOUCH) column is the unique spelling of the species in all data

```



```{r Join nutrient data with seafood database}
# Rename Seafood Species to Species
df_seafoodCompSimp_cleaned <- rename(df_seafoodCompSimp, Species = SeafoodSpecies)

names(df_seafoodCompSimp)[names(df_seafoodCompSimp) == "SeafoodSpecies"] <- "Species"

# Merge nutrient data with seafood database
df_seafood_compDB <- merge(df_UK_seafoodDB_capita, df_seafoodCompSimp, "Species")

str(df_seafood_compDB)

```







### Figure 5. Plot the top five seafoods that contribute the most to key daily RNIs (omega-3 fatty acids, calcium, iron, vitamin D and zinc) in UK adults. Stacked bar chart for each species

