---
title: "NatureFood_SeafoodFigures"
author: "Lofstedt, A."
date: "07-06-23"
output: powerpoint_presentation
---

Code for Nature Food figures only. 

### Preparation 

```{r Load libraries, include = FALSE}
# Load packages
library(tidyr) # for tidying data
library(dplyr) # base R alternative but neater
library(vroom) # for loading and transforming data
library(tidyverse) # data exploration 
library(reshape2) # for melting data frames i.e. short and wide to long and thin
library(data.table) # for fread()-function
library(mice) # md.pattern to show missing data
library(plyr) # joins multiple data frames together
library(ggplot2) # makes plots fancier
library(scales) # displays integers on the axis
library(ggthemes) # contains extra themes, scales and functions for and related to ggplot2
library(gridExtra)
library(cowplot) # Plots two figures one output
library(fmsb) # for radar plots

```


```{r Edit chunk settings, include = FALSE}
knitr::opts_chunk$set(
  echo = FALSE, 
  message = FALSE, 
  warninf = FALSE, 
  fig.height = 9, 
  fig.width = 16
)
```


```{r Setup default for figures, include=FALSE}
# Set default presentation of figures
theme_set(theme_bw()%+replace%
            theme(
              title = element_text(size = 26), 
              text = element_text(size = 30), 
              axis.text = element_text(size = 26, colour = "gray40"), # increase to make readable
              legend.key.size = unit(1.2, "cm"), 
              legend.title = element_text(size = 26), 
              legend.text = element_text(size = 24),
              panel.grid.major = element_line(colour= "gray40", linetype = "dotted", linewidth = 0.9), 
              panel.grid.minor = element_line(colour = "gray70", linetype = "dotted", linewidth = 0.7)
            ))
# instead of "size" can also use "linewidth"
```


```{r Set colour-blind palette}
# Colour-blind friendly palette with grey. Try to keep the order (http://www.cookbook-r.com/Graphs/Colors_(ggplot2)/#a-colorblind-friendly-palette)
cbPalette <- c( "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7", "#999999")
#consumption, exports, imports, production, purchases, supplies

# Colours will be allocated in commodity alphabetical order. 
```


# Add in per capita per week 


```{r Load file path, echo=TRUE}
# Loading file path to the data into a value 
source("Data_filepath.R")# Data_filepath.R is listed in .gitignore-file. So, you will need to create that file yourself and provide your respected filepath using "data_dir <- [enter your here]"
```

```{r Specify the desired file path}
filepath <- paste(data_dir,"Outputs/Databases", sep = "") 
# note we are able to change this to another file e.g. saving the output
```

```{r Load the seafood database}
# Load database 
df_UK_seafoodDB <- vroom(file=paste(filepath,"Seafood-database_preliminary.csv", sep="/"), na = c("NA"))

df_UK_seafoodDBNC <- vroom(file=paste(filepath,"Seafood-database-PCW_composition_preliminary.csv", sep="/"), na = c("NA"))
```



```{r Load population data}
#Specify the file path
filepath <- paste(data_dir,"ProcessedData", sep = "") 

# Load database 
df_UK_population <- vroom(file=paste(filepath,"UKPopulation_ONS_PreliminaryCleaned.csv", sep="/"), na = c("NA"))

# Simplify population data
df_UK_population_simp <- df_UK_population[c("Year", "UKPOP")]

```

```{r Merge SFDB with UK population}

df_UK_seafoodDB_Y <- merge(df_UK_seafoodDB, df_UK_population_simp, "Year")

```


```{r Changing volumne units to grams/capita/week}
# Calcualte g/capita/week
df_UK_seafoodDB_PCW <- 
  df_UK_seafoodDB_Y %>% 
  mutate(Volume = Volume/UKPOP/52, Units = "g/capita/week")

# Change structure of the database (short and wide)
# Aggregate value by commodity, year and species (production data composed of aquaculture and wild capture)
df_UK_seafoodDB_PCW_simp <- aggregate(Volume ~ Commodity + Year + Species + SpeciesType + SACN, data = df_UK_seafoodDB_PCW, sum)

# Need to unstack database to calculate seafood supplied. Change structure of the dataframe- convert to wide and short
df_UK_seafoodDB_PCW_short <- df_UK_seafoodDB_PCW_simp %>%
  pivot_wider(names_from = Commodity, values_from = Volume)

# Calculate seafood supplies whilst ignoring NA values
df_UK_seafoodDB_PCW_short <- df_UK_seafoodDB_PCW_short %>% 
  mutate(Supply = Production + Imports - Exports )
```



### Figure 1A. UK seafood supplies, consumption and purchases (g/capita/week) between 2009 and 2018.

```{r Seafood supplies consumption and purchases between 2009 and 2018}
# Subset desired variables 
df_UK_seafoodDB_PCW_shorter <- df_UK_seafoodDB_PCW_short[c("Supply", "Consumption", "Purchases", "Year", "Species", "SACN")]

# Change structure of the data set 
df_UK_seafoodDB_PCW_long <- melt(df_UK_seafoodDB_PCW_shorter, id.vars = c("Year", "Species", "SACN"), variable.name = "Commodity", value.name = "Volume")

# Aggregate data 
df_UK_seafoodDB_PCW_sum <- aggregate(Volume ~ Year + Commodity, data = df_UK_seafoodDB_PCW_long, sum)

# Plot total seafood supplied, purchases and consumed (g/capita/week) between 2009 and 2018. 
ggplot(df_UK_seafoodDB_PCW_sum, aes(Year, Volume, fill= Commodity)) +
  geom_bar(position = "dodge", stat= "identity", color = "black") +
  scale_y_continuous(name = "Amount of seafood (g/capita/week)", limits = c(0,300)) +
  scale_x_continuous(name = "Year", breaks = seq(2009, 2018, 1), 
                     guide = guide_axis(angle = 45)) +
  geom_hline(yintercept = 280, linetype = "dashed", color = "red", linewidth = 2) +
  scale_fill_manual(values = c("#009E73", "#E69F00", "#0072B2")) + # supplies, consumption, purchases
  theme(legend.title = element_blank()) 
  #theme(legend.position = "none")


```





### Figure 3. Total UK domestic UK seafood production (g/capita/week), bu species type between 2009 and 2018.

```{r Domestic seafood production}
# Subset production commodity 
df_UK_seafoodProduction_PCW <- df_UK_seafoodDB_PCW_short[c("Production", "Year", "Species", "SACN")]

# Sum UK seafood production, per SACN
df_UK_seafoodProduction_SACN <- aggregate(Production ~ SACN + Year, data = df_UK_seafoodProduction_PCW, sum)

# Select for lean and oily fish only
# Include "other" seafood types e.g. cephalopods and unknowns? 
df_UK_seafoodProduction_LOS <- subset(df_UK_seafoodProduction_SACN, SACN == "Lean" | SACN == "Oily" | SACN == "Shellfish")

df_UK_seafoodProduction_LOSO <- subset(df_UK_seafoodProduction_SACN, SACN == "Lean" | SACN == "Oily" | SACN == "Shellfish" 
                                        | SACN == "Other" )

# State order of the seafood types
df_UK_seafoodProduction_LOS$SACN_f <-  factor(df_UK_seafoodProduction_LOS$SACN, levels = c("Shellfish", "Lean", "Oily"))

df_UK_seafoodProduction_LOSO$SACN_f <-  factor(df_UK_seafoodProduction_LOSO$SACN, levels = c("Other", "Shellfish", 
                                                                                             "Lean", "Oily"))


# Plot average seafood supply (g/capita/week), to SACN level (lean and oily) between 2009 and 2018. 
ggplot(df_UK_seafoodProduction_LOSO, aes(y = Production, x = Year, fill = SACN_f)) +
  geom_bar(position = "stack", stat = "identity", color = "black", width = 0.5) + 
  scale_y_continuous(name = "Total seafood production (g/capita/week)") +
  scale_x_continuous(name = "Year", breaks = seq(2009, 2018, 1), 
                     guide = guide_axis(angle = 45)) +
  scale_fill_manual(values = c("#999999", "#DDCC77",  "#CC79A7","#D55E00")) +
  geom_hline(yintercept = 280, linetype = "dashed", color = "red", linewidth = 2.0)

```




### Figure 5. Species contibutions to RNIs Nutrient supply, by species, in 2018. 

Key nutrients include calcium, iron, vitamin D, vitamin A, vitamin B12, iodine, selenium, omega-3 fatty acids and zinc.

# Create radar charts  


```{r Calculate nutrient supplies from seafood  nutrient value by supplies}
# Multiply nutrient value (g/100g) bu seafood supplies (g/capita/day) and divide by 100. 

```


```{r Join data frame with translation document}
# Cleaned data set does not consider the status of the sample e.g. raw/ cooked. So new column created with such information.
# Duplicate species name column (duplicated column entries to be duplicated)
#df_seafoodCompSimp$SpeciesRevised <- df_seafoodCompSimp$SampleDescription

# Merge data frame with MCS data frame by species name
#df_seafoodCompSimp <- df_seafoodCompSimp %>% inner_join(., MCS_translations, by = c('SpeciesRevised' = 'MCS (DO NOT TOUCH)'))

# Check the structure
#str(df_seafoodCompSimp)
```




```{r Calculate nutrient density score of each fish species}

# Divide nutrient (grams per 100 grams of food) by the recommended daily allowance (RDA) for that nutrient.

# Then divide calories (per 100 grams of the food) by the reference calorie intake (2,000 calories per day).

# Divide the RDA calculation by the calorie calculation 



```{r filter data for figure 4}
df_figure_input <- df_SFDB_PCW_composition %>%  
  filter(Year == "2018", !SACN == "Other", !is.na(Production), !is.na(Imports), !is.na(Exports), !is.na(VitaminD_µg100g), !is.na(Zinc_mg100g), !is.na(TotalPUFA_g100g), !is.na(Calcium_mg100g), !is.na(Iron_mg100g)) %>% 
  mutate_if(is.character, factor) %>% 
  select(Species, Year, SpeciesType, SACN, Production, Exports, Imports, UKPOP, TotalPUFA_g100g, Calcium_mg100g, Iron_mg100g, VitaminD_µg100g, Zinc_mg100g)


```


```{r Calculation for Figure 4}

df_figure_input <- df_figure_input %>% 
  mutate(Supply = Production + Imports - Exports ) %>% 
  mutate(Supply = Supply/UKPOP/52) %>% # g/capita/week
  relocate(Supply, .after= SACN) %>% 
  select(-Production, - Imports, - Exports, - UKPOP)

```


omega-3 fatty acids, calcium, iron, vitamin D and zinc

```{r Radar Plot}
# Load nutrient composition data
df_UK_seafoodDBNC 

df_UK_NC_SOI <- subset(df_UK_seafoodDBNC, Species == "Herring" | Species == "Mackerel"| Species == "Salmon")

# Select desired nutrients
df_UK_NC_NOI <- df_UK_NC_SOI %>%  
  select(Species, TotalPUFA_g100g, Calcium_mg100g, Iron_mg100g, VitaminD_µg100g, Zinc_mg100g) %>%
  group_by(Species) %>%
  filter(row_number()==1)



## up to here

# Data needs to be structured in a certain way. First row: maximum values for each variable; 
# Second row: minimum values for each variable, Third row: Data for variables  

# Define the variable ranges: maximum and minimum for each nutrient. Set max to recommended RNI 
max_min <- data.frame(
  PUFA = c(20, 0), calcium = c(20, 0), Maths = c(20, 0),
  Sport = c(20, 0), English = c(20, 0), Geography = c(20, 0),
  Art = c(20, 0), Programming = c(20, 0), Music = c(20, 0)
) # change subject to nutrients
rownames(max_min) <- c("Max", "Min") # State row names

# Plot the data for one fish 
student1_data <- df[c("Max", "Min", "Student.1"), ]
radarchart(student1_data)

# Plot the data for three fish in the same plot

# Reduce plot margin using par()
op <- par(mar = c(1, 2, 2, 2))

# Create the radar charts
create_beautiful_radarchart(
  data = df, caxislabels = c(0, 5, 10, 15, 20), # state labels- change to nutrient values
  color = c("#00AFBB", "#E7B800", "#FC4E07") # States colors
)
# Add an horizontal legend 
legend(
  x = "bottom", legend = rownames(df[-c(1,2),]), horiz = TRUE,
  bty = "n", pch = 20 , col = c("#00AFBB", "#E7B800", "#FC4E07"),
  text.col = "black", cex = 1, pt.cex = 1.5
  )
par(op)
```


```{r Seperate radar plots}
# Define colors and titles
colors <- c("#00AFBB", "#E7B800", "#FC4E07")
titles <- c("Student.1", "Student.2", "Student.3") # change to fish 

# Reduce plot margin using par()
# Split the screen in 3 parts
op <- par(mar = c(1, 1, 1, 1))
par(mfrow = c(1,3))

# Create the radar chart
for(i in 1:3){
  create_beautiful_radarchart(
    data = df[c(1, 2, i+2), ], caxislabels = c(0, 5, 10, 15, 20),
    color = colors[i], title = titles[i]
    )
}
par(op)


# Or ggradar plot
# Plot for one student
ggradar(
  df[1, ], 
  values.radar = c("0", "10", "20"),
  grid.min = 0, grid.mid = 10, grid.max = 20
  )
```


https://www.datanovia.com/en/blog/beautiful-radar-chart-in-r-using-fmsb-and-ggplot-packages/#:~:text=A%20radar%20chart%2C%20also%20known%20as%20a%20spider,packages%3A%20the%20fmsb%20or%20the%20ggradar%20R%20packages.



### Figure 1B. UK seafood supplies and consumption (g/capita/week), by species, in 2018

```{r Seafood supplies and consumption between 2009 and 2018}
# Subset desired variables 
df_UK_seafoodDB_PCW_CS <- df_UK_seafoodDB_PCW_short[c("Supply", "Consumption", "Year", "Species", "SACN")]

# Change structure of the data set 
df_UK_seafoodDB_PCW_CS_long <- melt(df_UK_seafoodDB_PCW_CS, id.vars = c("Year", "Species", "SACN"), variable.name = "Commodity", value.name = "Volume")

# Aggregate data 
df_UK_seafoodDB_PCW_CS_sum <- aggregate(Volume ~ Year + Commodity, data = df_UK_seafoodDB_PCW_CS_long, sum)

# Plot total seafood supplied and consumed (g/capita/week) between 2009 and 2018. 
ggplot(df_UK_seafoodDB_PCW_CS_sum, aes(Year, Volume, fill= Commodity)) +
  geom_bar(position = "dodge", stat= "identity", color = "black") +
  scale_y_continuous(name = "Amount of seafood (g/capita/week)", limits = c(0,300)) +
  scale_x_discrete(guide = guide_axis(angle = 45)) +
  geom_hline(yintercept = 280, linetype = "dashed", color = "red", linewidth = 2) +
  scale_fill_manual(values = c("#009E73", "#E69F00")) + # supplies, consumption, 
  theme(legend.title = element_blank()) +
  theme(legend.position = "none")

```


